### 2024-08-06 doc/paper.qmd
finalise documentation

---
title: |
  | Earnings Management and Investor Protection:
  | Accounting Reading Group - Assignment III\vspace{1cm}
author:
  - name: Melisa Mazaeva
    email: melisa.mazaeva@student.hu-berlin.de
    affiliations:
      - Humboldt-Universität zu Berlin  
date: today
date-format: MMM D, YYYY [\vspace{1cm}]
abstract: |
  | This project uses the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) to provide an infrastructure for open science-oriented empirical projects. Leveraging external Worldscope data sets on financial data, the repository showcases a reproducible workflow that integrates Python scripts for data analysis. The project’s output demonstrates a comprehensive application of skills to replicate and extend the findings from the seminal paper by Leuz, Nanda, and Wysocki (2003), particularly in providing descriptive statistics for the four individual earnings management measures as well as the aggregate earnings management score across various countries. In doing so, it documents and discusses the research design choices made and the variations between the original and reproduced results. This code base, adapted from TREAT, should give you an overview on how the template is supposed to be used for my specific project and how to structure a reproducible empirical project.
  | \vspace{6cm}
bibliography: references.bib
biblio-style: apsr
format:
  pdf:
    documentclass: article
    number-sections: true
    toc: false
fig_caption: yes
fontsize: 11pt
ident: yes
always_allow_html: yes
header-includes:
  - \usepackage[nolists]{endfloat}    
  - \usepackage{setspace}\doublespacing
  - \setlength{\parindent}{4em}
  - \setlength{\parskip}{0em}
  - \usepackage[hang,flushmargin]{footmisc}
  - \usepackage{caption} 
  - \captionsetup[table]{skip=24pt,font=bf}
  - \usepackage{array}
  - \usepackage{threeparttable}
  - \usepackage{adjustbox}
  - \usepackage{graphicx}
  - \usepackage{csquotes}
  - \usepackage{indentfirst}  # Added this line to ensure the first paragraph is indented for better readability
  - \usepackage[margin=1in]{geometry}
---

\pagebreak

## delete or take fikirs below
```{python}
#| echo: false
#| output: false

import pickle

with open('../output/results.pickle', 'rb') as f:
        results = pickle.load(f)

min_fyear, max_fyear, no_unique_firms = results['Desc information'].values()

def rename_var_to_label(table, italic=True):
    for var_name, var_label in results['Variable Names'].items():
        var_label = f'\\textit{{{var_label}}}' if italic else var_label
        table = table.replace(var_name, var_label)
    return table
```

## from fikir
```{python}
#| echo: false
#| output: false

import pickle
import pandas as pd

with open('../output/results.pickle', 'rb') as f:
        results = pickle.load(f)

def escape_underscores(df):
    df.columns = [col.replace('_', '\_') for col in df.columns]
    return df

# Custom formatter function
def custom_float_format(x):
    return f"{x:.3f}".lstrip('0') if pd.notnull(x) else ""

def prep_latex_table(df, caption=None, label=None):
    df = escape_underscores(df.reset_index())
    num_columns = len(df.columns)
    column_format = 'l' + 'r' * (num_columns - 1)
    latex_table = df.to_latex(
        column_format=column_format,
        index=False,
        float_format=custom_float_format,
    )
    latex_table_lines = [
        "\\begin{table}[htbp]",
        "\\centering",
        "\\resizebox{\\textwidth}{!}{%",
        latex_table,
        "}",
        f"\\caption{{{caption}}}" if caption else "",
        f"\\label{{{label}}}" if label else "",
        "\\end{table}"
    ]
    return "\n".join(line for line in latex_table_lines if line)
```

\pagebreak

## delete. from mine
# List of Abbreviations
\setlength{\parindent}{0em}

**EM**: Earnings Management
**IDE**: Integrated Development Environment  
**UK**: United Kingdom  
**US**: United States  
**WRDS**: Wharton Research Data Services

\setlength{\parindent}{4em}

\pagebreak

# Introduction

The following paper illustrates how to use open science tools in empirical accounting research.

The original study by @Leuz_2003 examines systematic differences in earnings management (EM) across 31 countries and proposes that these differences are influenced by the level of investor protection. Specifically, they argue that stronger investor protection reduces earnings management by limiting insiders’ ability to acquire private control benefits.

This paper presents details on a replication of Table 2 Panel A from the seminal paper by @Leuz_2003. The original table provides descriptive statistics for four individual earnings management measures and the aggregate earnings management score across various countries, highlighting significant differences in EM practices between Continental Europe, Asia, and Anglo-American countries.

The project uses data from the Worldscope Database through Wharton Research Data Services (WRDS), focusing on the fiscal years 1990 to 1999 and covering multiple countries. The replication aims to follow the research design by @Leuz_2003 as closely as possible with the available data. It employs four individual earnings management measures: earnings smoothing using accruals (EM1), the correlation between changes in accruals and cash flows (EM2), the magnitude of accruals (EM3), and the ratio of small profits to small losses (EM4). These measures are combined to create an aggregate earnings management score, providing a comprehensive overview of EM practices.

The analysis includes pulling and filtering the relevant financial data, identifying relevant companies and countries based on the authors’ requirements, calculating the earnings management measures for each country, and examining the relationship between each country’s aggregate earnings management score and its ranking based on made explicit assumptions whenever the paper is not clear on how to proceed.

## delete. from mine
Note that intermediate check steps have been omitted for brevity. The full code, including intermediate print steps to verify outputs, is presented in `code/python/Assignment1_analyze_residual_income.py`.

The replicated table is shown below, followed by a detailed comparison, replication steps documentation, and discussion of the results.

## delete from fikir. 1st table is for original table
```{python}
#| echo: false
#| output: asis

print(prep_latex_table(
    results['penman_table'],
    caption='Original table Penman 2013 - P/B Ratios And Future Residual Income',
    label='tab:original_penman'
))
```

```{python}
#| echo: false
#| output: asis

print(prep_latex_table(
    results['replication_results'],
    caption='Replicated table - P/B Ratios And Future Residual Income',
    label='tab:replicated_penman'
))
```

# Research Design Choices and Assumptions

The aim of Assignment III is to replicate a specific empirical table (Table 2 Panel A) from the seminal paper by @Leuz_2003. This table involves calculating the EM measures for firms across various countries over a defined period and examining the relationship between these measures and investor protection. The replication process includes data loading, preparation, cleaning, and normalization, followed by the application of statistical methods to compute and interpret financial metrics. For Assignment III, I pulled data from the Worldscope database through WRDS and used the Python programming language to carry out the empirical analysis. Visual Studio Code was used as the Integrated Development Environment (IDE) for writing, debugging, and optimizing the Python code.

The replication is based on data pulled from the Worldscope database, specifically from the `wrds_ws_company` and `wrds_ws_funda` tables, which were merged for the analysis. The first table provides company profile information, including items such as ISIN, Worldscope Identifiers, company name, and the country where the company is domiciled @WRDS_WS_Company_2024. The latter table contains Fundamentals Annual data at the company-year level, including items such total assets, net income, and other relevant financial variables @WRDS_WS_Funda_2024.

Following @Leuz_2003, I focus the analysis on companies across various countries, ensuring that the data accurately reflects the fiscal years 1990 to 1999 as specified in the original study. The replication aims to mirror the research design as closely as possible with the available data.

In addition, I impose the following assumptions to ensure clarity and consistency where the paper does not provide explicit guidance:

1. The original paper references the November 2000 version of the Worldscope Database. However, the data used for this analysis represents the latest available version, updated in July 2024, with quarterly frequency updates [@WRDS_2024]. Due to potential adjustments and updates made to the database since November 2000, there may be differences between the databases that could affect the results. For example, companies may restate financials after the original reporting period, so that these restatements are reflected in the later database version rather than the historical one. Moreover, the data vendor Refinitiv regularly updates its databases to correct errors and add new information, which may be included in the later data but not in the November 2000 snapshot.
2. The original paper outlines key terms that will be used in this project to ensure consistency and accuracy in the replication. @Leuz_2003 define earnings management as the manipulation of a firm's reported economic performance by insiders to deceive certain stakeholders or to affect contractual results. Authors describe investor protection as a key institutional factor that limits insiders' acquisition of private control benefits, thereby reducing their incentives to manage accounting earnings by ensuring strong and well-enforced rights for outside investors. Finally, private control benefits are the benefits that insiders can gain from controlling a firm, which can include financial gains or other advantages that are not shared with other stakeholders @Leuz_2003.
3. While pulling the data for analysis, I encountered negative values for some key financial metrics such as operating income (`item1250`), net income before extraordinary items/preferred dividends (`item1551`), and net income before preferred dividends (`item1651`). The paper by @Leuz_2003 does not explicitly specify how to handle negative values in key financial metrics. For the purpose of this replication study, I will include negative values in the analysis. Including these values ensures that the analysis captures the full spectrum of earnings management activities across different countries.
4. Another potential source of discrepancies between the original and replicated tables may be the choice of variables pulled from Worldscope. For example, in the `wrds_ws_funda table`, both `item1151` and `item4051` are named “DEPRECIATION, DEPLETION AND AMORTIZATION” @WRDS_WS_Funda_2024. I chose `item1151` from the Income Statement rather than `item4051` from the Cash Flow Statement, based on the Excel Industrials Template by @WRDS_Balancing_Model_2024. Since the authors do not specify the choice of variables used from the database, this could cause differences in the results.
5. The EM measures are based on scaled variables (e.g., operating cash flow scaled by lagged total assets). As such, the currency of the relevant data items should not affect the results as long as the same currency is used consistently for both the numerator and the denominator. This approach ensures comparability across different countries, regardless of their local currencies. Additionally, according to a document by [@Thomson_Financial_2007, p.20], all Worldscope data is consistently reported in the local currency of each firm’s country of domicile, eliminating the need for currency conversions in this project.
333. ## delete. mine 
333. It is assumed that Penman (2013) restricted the possible P/B values to a range of 0 to 7 to exclude outliers. Hence, extreme P/B values that are negative and very high (greater than 7) are excluded to focus the analysis on firms with more stable and reasonable valuations, reducing the impact of outliers.

By following the steps provided in Section 4 and adhering to the assumptions made, I successfully replicated the analysis and produced the required table. A thorough step-by-step approach, with each step clearly documented,  helped to understand and verify the outputs.

# Replication Steps
## Step 1: Pulling the Data and Managing the Databases
In contrast to Assignment I, where the data was provided externally, Assignment III involves additionally pulling data directly from the Worldscope database, merging relevant tables, and preparing the data for further analysis from raw data to final output.

To compile the dataset, the dynamic and static datasets were merged on the `item6105`  identifier, representing the unique Worldscope Permanent ID. WRDS advises using this identifier consistently within Worldscope data since it remains stable over time [@WRDS_Overview_2024]. An inner join was used for this merge, because this approach ensures that only the complete and consistent data from both tables is retained.

Moreover, an additional filter is applied to retain only company rows. This is achieved by selecting rows where the `item6100` field equals 'C', indicating that the Worldscope Identifier represents a company. This step ensures that the analysis includes only company data, excluding averages, exchange rates, securities, or stock indices, as indicated by [@WRDS_Overview_2024].

As required by @Leuz_2003, financial institutions are excluded from the analysis based on their SIC codes. This is done by filtering out rows where the `item7021` identifier, representing the SIC code, falls within the range of 6000 to 6999. Hence, the dataset focuses only on non-financial companies, aligning with the methodology of the original paper.

Finally, additional filtering ensures that only data from the 31 countries, as given in the paper, is included. These filtering steps were applied to reduce the dataset size and improve the workflow.

Notably, the configuration file utilizes additional refinement as proposed by [@WRDS_Overview_2024]. Using `A` (Annual) in `freq` variable ensures that the data represents the financial information reported on an annual basis, which is consistent with the paper’s methodology. This excludes data reported on a current, daily, or quarterly basis.

After retaining only relevant company data, excluding financial institutions, and focusing on specified countries, the processed data was then saved to a CSV file at the path `data/pulled` specified in the configuration file `config/pull_data_cfg​`.

## Step 2: Data Preparation
To prepare the data according to the methodology outlined in the paper, it is essential to follow specific steps. Firstly, countries with sufficient firm-year observations must be filtered. Each country should have at least 300 firm-year observations for certain key accounting variables, including total assets, sales, net income, and operating income. Secondly, firms with adequate consecutive data must be identified. Each firm must have income statement and balance sheet information for at least three consecutive years, where I focus on the same key variables as in the previous step.

## Step 3: Analysis Implementation and Table Reproduction
In this replication step, implement the calculations for the EM measures as described in the methodology of the paper, ensuring that all statistical methods and groupings are accurately replicated.

# Discretionary Accruals ##delete?

To demonstrate our workflow, we explore discretionary accruals across the U.S. We calculate modified Jones and Dechow and Dichev type accruals and show their distributional properties. The main purpose of all this, however, is to provide a toy use case for [our project template directory](https://github.com/trr266/project_template) that contains all the code to obtain the data, run the analysis and prepare a paper as well as a presentation.

Table \ref{tab:tab-descriptives-sample} presents our data that is based on a simple WRDS pull of Compustat data with financial firms (SIC 6XXX) excluded. We require data to calculate all variables and this drastically reduces the sample size. Modified Jones discretionary accruals are calculated loosely based on @HN_2007 and Dechow and Dichev discretionary accruals are calculated based on (big surprise) @DD_2002. As you will see from @fig-boxplot_smp, discretionary accruals are very noisy constructs, even after limiting the sample to observations with complete data and winsorizing all data to the top and bottom percentile for each year. @fig-scatter_dd_lnta shows a very prominent heteroscedasticity of discretionary accruals with regards to size. While researchers have tried to address this problem, the distributional properties of these constructs significantly complicate the interpretation of discretionary accrual-related findings. Especially in high powered settings, the measurement error, being highly correlated with size, will tend to load on variables that are unrelated to the underlying economic construct but correlated with size. Table \ref{tab:corr} shows some correlations and Table \ref{tab:reg} shows some completely pointless regressions.

## delete. cell 3 on Descriptive Statistics Table
```{python}
#| label: fig-boxplot_smp
#| fig-cap: "Distribution of Discretionary Accruals over Time"
#| echo: false
#| fig-format: svg

results['fig_boxplot_smp']
```

```{python}
#| label: fig-scatter_dd_lnta
#| fig-cap: "Dechow and Dichev DA and Firm Size"
#| echo: false
#| fig-format: svg

results['fig_scatter_dd_lnta']
```

```{python}
#| echo: false
#| output: asis

lat_tab = rename_var_to_label(results['Descriptive Statistics']).split('\n')
lat_tab[4] = '\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]'
lat_tab[6] = '\\hline\\\\[-1.8ex]'
lat_tab[-4] = '\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]'

print('\n'.join([
    '\\begin{table}[!htbp]',  
    '\caption{\label{tab:descriptives}Descriptive Statistics}',
    '\label{tab:tab-descriptives-sample}',
    '\\begin{adjustbox}{width=\\textwidth}',
    '\\begin{threeparttable}',
    '\n'.join(lat_tab[3:-2]),
    '\\begin{tablenotes}[flushleft]',
    '\setlength{\labelsep}{0pt}',
    '\\footnotetext',
    f'\\item Note: The data is obtained from the Compustat U.S. as provided by WRDS. The sample covers the period {int(min_fyear)} to {int(max_fyear)} and {no_unique_firms:,} unique firms.',
    '\end{tablenotes}',
    '\end{threeparttable}',
    '\end{adjustbox}',
    '\\end{table}',
]))
```

```{python}
#| echo: false
#| output: asis

lat_tab = rename_var_to_label(results['Correlations'], False).split('\n')

lat_tab[2] = '\\\\[-0.9ex]\\hline \\hline \\\\[-0.9ex]'
lat_tab[4] = '\\hline\\\\[-1.8ex]'
lat_tab[12] = '\\\\[-0.9ex]\\hline \\hline \\\\[-0.9ex]'
lat_tab[14] = '\\begin{tablenotes}[flushleft]'

print('\n'.join([
    "\\begin{table}[!htbp]",
    "\centering",
    "\caption{\label{tab:CorrTable}Correlations}",
    '\label{tab:corr}',
    '\n'.join(lat_tab),
    "\end{table}"
]))
```

```{python}
#| echo: false
#| output: asis

print('\n'.join([
    '\\begin{table}[!htbp] \\centering',
    '\\caption{\\label{tab:reg}Regressions}', 
    '\label{tab:reg}',
    rename_var_to_label('\n'.join(results['Regression'].split('\n')[3:-1])),
    "\end{table}"
]))
```

# Results
Cross-check your results with the original table to ensure accuracy. Compare your results with the results in the paper. If you notice differences, discuss where they may originate. Make sure to be as specific as possible. This could involve looking at the data selection, calculation methods, or any assumptions that might differ from the original; changes in the data over time, different interpretations of the methodology, or limitations in the data sources used. Consider whether the differences are significant and what they might mean for the conclusions drawn in the paper.

# Conclusion

Isn't that wonderful? Discretionary accruals rock but what rocks even more is open science and a collaborative workflow. Clone or fork this repository to kickstart your own projects. If you do not like either R or Python, consider contributing code in your favorite statistical programming language to the repo. Thanks for reading and enjoy!
gpt: Summarize the key findings and their significance.
This final assignment has required the comprehensive application of all skills learned and feedback received from previous assignments, making it a more complex project workflow.

The study by @Leuz_2003 documents systematic differences in the level of earnings management across 31 countries. It suggests that insiders use EM to conceal firm performance and protect their private control benefits and that stronger investor protection reduces earnings management by limiting insiders’ ability to gain private benefits. By conducting a descriptive cluster analysis after descriptive statistics that was replicated in this project, the paper identifies groupings of countries with similar institutional characteristics and demonstrates that earnings management varies systematically across these institutional clusters.

\pagebreak

\setcounter{table}{0}
\renewcommand{\thetable}{\arabic{table}}

# References {-}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\noindent

### 2024-08-06 code/python/pull_wrds_data.py

@@ -68,18 +68,18 @@ def pull_wrds_data(cfg, wrds_authentication):
     db.close()
     log.info("Disconnected from WRDS")
 
-    # Merge the static and dynamic data to create one single data set by the unique ITEM6105 (Worldscope Permanent ID)
+    # Merge the static and dynamic data to create one single data set by the unique item6105 (Worldscope Permanent ID)
     wrds_data = pd.merge(wrds_data_dynamic, wrds_data_static, left_on='item6105', right_on='item6105', how='inner')
 
     # Apply the filter for entity type to select only company rows in merged data based on Worldscope-specific Identifier advice
-    # to retrieve only company and drop average, exchange rate, ADR, security or a stock index.
+    # to retrieve only companies and drop average, exchange rate, ADR, security or a stock index.
     wrds_data = wrds_data[wrds_data['item6100'] == 'C']
 
     # Filter out financial institutions using SIC code as required by paper
     wrds_data = wrds_data.dropna(subset=['item7021'])
     wrds_data = wrds_data[(wrds_data['item7021'].astype(int) < 6000) | (wrds_data['item7021'].astype(int) > 6999)]
 
-    # Apply the filter for specified countries
+    # Apply the filter for specified countries given in the paper
     wrds_data = wrds_data[wrds_data['item6026'].isin(cfg['included_countries']) & ~wrds_data['item6026'].isin(cfg['excluded_countries'])]
 
 

### 2024-08-07 doc/paper.qmd
pulled and prepared data

@@
 
-## delete. from mine
 # List of Abbreviations
@@
 
-The replicated table is shown below, followed by a detailed comparison, replication steps documentation, and discussion of the results.
+The replicated table is shown below, followed by replication steps documentation, a detailed comparison, and discussion of the results.
 
-## delete from fikir. 1st table is for original table
-```{python}
-#| echo: false
-#| output: asis
-
-print(prep_latex_table(
-    results['penman_table'],
-    caption='Original table Penman 2013 - P/B Ratios And Future Residual Income',
-    label='tab:original_penman'
-))
-```
-
+## delete from fikir.
 ```{python}
@@
 
-In addition, I impose the following assumptions to ensure clarity and consistency where the paper does not provide explicit guidance:
+In addition, I impose the following assumptions to ensure clarity and consistency where the paper by @Leuz_2003 does not provide explicit guidance:
 
@@
 3.
-While pulling the data for analysis, I encountered negative values for some key financial metrics such as operating income (`item1250`), net income before extraordinary items/preferred dividends (`item1551`), and net income before preferred dividends (`item1651`).
+While pulling the data for analysis, I encountered negative values for some key financial metrics such as operating income (`item1250`), or net income before preferred dividends (`item1651`).
 The paper by @Leuz_2003 does not explicitly specify how to handle negative values in key financial metrics.
@@
 
+To ensure the integrity of the data, I filtered out rows with empty `item6105` (Worldscope Permanent ID) values, as this identifier is critical for firm/year level filtering in the data preparation step.
+In total, 125 observations from the dynamic data and 10,306 observations from the static data were removed.
 To compile the dataset, the dynamic and static datasets were merged on the `item6105` identifier, representing the unique Worldscope Permanent ID.
@@
 
-As required by @Leuz_2003, financial institutions are excluded from the analysis based on their SIC codes.
+As required by @Leuz_2003, financial institutions are removed from the analysis based on their SIC codes.
 This is done by filtering out rows where the `item7021` identifier, representing the SIC code, falls within the range of 6000 to 6999.
@@
 
-After retaining only relevant company data, excluding financial institutions, and focusing on specified countries, the processed data was then saved to a CSV file at the path `data/pulled` specified in the configuration file `config/pull_data_cfg​`.
+After retaining only relevant company data, filtering out financial institutions, and focusing on specified countries, the processed data was then saved to a CSV file at the path `data/pulled` specified in the configuration file `config/pull_data_cfg​`.
 
 ## Step 2: Data Preparation
-To prepare the data according to the methodology outlined in the paper, it is essential to follow specific steps.
+First, I checked the dataset for duplicate firm-year observations based on the combination of the Worldscope Permanent ID (`item6105`) and the fiscal year (`year_`) and confirmed that there were no duplicates present, ensuring the accuracy of the data for further analysis.
+As the original study does not specify the net income measure, `item1651` (Net Income before Preferred Dividends) was used as the variable for net income.
+This choice aligns with the final net income figure reported in the income statement, based on the Excel Industrials Template by @WRDS_Balancing_Model_2024.
+
+To prepare the data sample in line with the methodology outlined in the paper, it is essential to follow the requirements, as instructed by @Leuz_2003.
 Firstly, countries with sufficient firm-year observations must be filtered.
@@
 Each firm must have income statement and balance sheet information for at least three consecutive years, where I focus on the same key variables as in the previous step.
+If a firm had at least three consecutive years of complete data at any point, all its data entries were retained in the final dataset.
 
@@
 In this replication step, implement the calculations for the EM measures as described in the methodology of the paper, ensuring that all statistical methods and groupings are accurately replicated.
-
-# Discretionary Accruals ##delete?
-
-To demonstrate our workflow, we explore discretionary accruals across the U.S.
-We calculate modified Jones and Dechow and Dichev type accruals and show their distributional properties.
-The main purpose of all this, however, is to provide a toy use case for [our project template directory](https://github.com/trr266/project_template) that contains all the code to obtain the data, run the analysis and prepare a paper as well as a presentation.
-
-Table \ref{tab:tab-descriptives-sample} presents our data that is based on a simple WRDS pull of Compustat data with financial firms (SIC 6XXX) excluded.
-We require data to calculate all variables and this drastically reduces the sample size.
-Modified Jones discretionary accruals are calculated loosely based on @HN_2007 and Dechow and Dichev discretionary accruals are calculated based on (big surprise) @DD_2002.
-As you will see from @fig-boxplot_smp, discretionary accruals are very noisy constructs, even after limiting the sample to observations with complete data and winsorizing all data to the top and bottom percentile for each year.
-@fig-scatter_dd_lnta shows a very prominent heteroscedasticity of discretionary accruals with regards to size.
-While researchers have tried to address this problem, the distributional properties of these constructs significantly complicate the interpretation of discretionary accrual-related findings.
-Especially in high powered settings, the measurement error, being highly correlated with size, will tend to load on variables that are unrelated to the underlying economic construct but correlated with size.
-Table \ref{tab:corr} shows some correlations and Table \ref{tab:reg} shows some completely pointless regressions.
 
@@
 # Results
-Cross-check your results with the original table to ensure accuracy.
+##Cross-check your results with the original table to ensure accuracy.
 Compare your results with the results in the paper.
@@
 
+The replicated table shows some similarities with the original table by @Leuz_2003 in that . These findings are in line with valuation theory that .
+
+However, the patterns are less clear in the replicated table (in particular ), which stands in contrast to @Leuz_2003.
+
 # Conclusion
 
-Isn't that wonderful?
-Discretionary accruals rock but what rocks even more is open science and a collaborative workflow.
-Clone or fork this repository to kickstart your own projects.
-If you do not like either R or Python, consider contributing code in your favorite statistical programming language to the repo.
-Thanks for reading and enjoy!
-gpt: Summarize the key findings and their significance.
-This final assignment has required the comprehensive application of all skills learned and feedback received from previous assignments, making it a more complex project workflow.
+##gpt: Summarize the key findings and their significance. take results and insert them here
+
+This project demonstrates the effectiveness of using an open science and collaborative workflow for analyzing earnings management measures across various countries.
+By following a step-by-step approach and using the TRR 266 Template for Reproducible Empirical Accounting Research, I was able to replicate an empirical table from @Leuz_2003 and provide insights into the trends of earnings management practices, which vary significantly across countries, similar to @Leuz_2003.
 
@@
 By conducting a descriptive cluster analysis after descriptive statistics that was replicated in this project, the paper identifies groupings of countries with similar institutional characteristics and demonstrates that earnings management varies systematically across these institutional clusters.
+
+The analysis revealed that
+
+The replication yields mixed results. While
+
+This final assignment has required the comprehensive application of all skills learned and feedback received from previous assignments, making it a more complex project workflow.
+In the future, this repository can be cloned or forked (if made public) to kickstart further projects on earnings management measures analysis.
+Thanks for reading!
 

### 2024-08-07 code/python/prepare_data.py

@@ -1,13 +1,10 @@
 # --- Header -------------------------------------------------------------------
-# Prepares the data for analysis
+# Prepare the pulled data for further analysis as per the requirements by Leuz et al.
 #
-# (C) Your Project -  See LICENSE file for details
+# (C) Melisa Mazaeva -  See LICENSE file for details
 # ------------------------------------------------------------------------------
 
-from itertools import product
-import numpy as np
 import pandas as pd
-
 from utils import read_config, setup_logging
 
 log = setup_logging()
@@ -17,53 +14,43 @@ def main():
     cfg = read_config('config/prepare_data_cfg.yaml')
 
     # Load the pulled data
-    data = pd.read_csv(cfg['worldscope_sample_save_path'])
-    log.info(f"Loaded data with {data.shape[0]} rows and {data.shape[1]} columns.")
-    log.info(f"Data columns: {data.columns.tolist()}")
+    wrds_data = pd.read_csv(cfg['worldscope_sample_save_path'])
 
-    # Filter countries with at least 300 firm-year observations for the required variables
-    required_vars = ['item2999', 'item1001', 'item7250', 'item1250']
+    # Check for duplicate firm-year observations
+    dup_obs = wrds_data[wrds_data.duplicated(subset=['item6105', 'year_'], keep=False)]
+    if not dup_obs.empty:
+        log.warning(f"Found {dup_obs.shape[0]} duplicate firm-year observations. Removing duplicates.")
+        wrds_data = wrds_data.drop_duplicates(subset=['item6105', 'year_'], keep='first')
 
-    # Filter out rows with missing values in the required variables
-    data_filtered = data.dropna(subset=required_vars)
-    log.info(f"Data after filtering missing values in required variables has {data_filtered.shape[0]} rows.")
+    # Filter countries with at least 300 firm-year observations for key accounting variables
+    filtered_countries_data = filter_countries(wrds_data)
 
-    # Count firm-year observations per country
-    country_counts = data_filtered['item6026'].value_counts()
-    log.info(f"Country counts:\n{country_counts}")
+    # Filter firms with at least three consecutive years of income statement and balance sheet information
+    filtered_firms_data = filter_firms(filtered_countries_data)
 
-    # Filter countries with at least 300 observations
-    sufficient_data_countries = country_counts[country_counts >= 300].index.tolist()
[…]

### 2024-08-07 code/python/pull_wrds_data.py

@@ -68,8 +68,17 @@ def pull_wrds_data(cfg, wrds_authentication):
     db.close()
     log.info("Disconnected from WRDS")
 
+    # Display the number of rows with empty item6105 before filtering
+    log.info(f"Rows with empty item6105 in dynamic data before filtering: {wrds_data_dynamic['item6105'].isna().sum()}")
+    log.info(f"Rows with empty item6105 in static data before filtering: {wrds_data_static['item6105'].isna().sum()}")
+
+    # Filter out rows with empty item6105 in both dynamic and static datasets, because this unique identifier is crucial 
+    # for further data preparation step when we filter on firm/year observations
+    wrds_data_dynamic = wrds_data_dynamic.dropna(subset=['item6105'])
+    wrds_data_static = wrds_data_static.dropna(subset=['item6105'])
+
     # Merge the static and dynamic data to create one single data set by the unique item6105 (Worldscope Permanent ID)
-    wrds_data = pd.merge(wrds_data_dynamic, wrds_data_static, left_on='item6105', right_on='item6105', how='inner')
+    wrds_data = pd.merge(wrds_data_dynamic, wrds_data_static, on='item6105', how='inner')
 
     # Apply the filter for entity type to select only company rows in merged data based on Worldscope-specific Identifier advice
     # to retrieve only companies and drop average, exchange rate, ADR, security or a stock index.

### 2024-08-07 doc/paper.qmd
modified preparation code

@@
 ## Step 2: Data Preparation
-First, I checked the dataset for duplicate firm-year observations based on the combination of the Worldscope Permanent ID (`item6105`) and the fiscal year (`year_`) and confirmed that there were no duplicates present, ensuring the accuracy of the data for further analysis.
+In order to verify the pulle data, I checked the dataset for duplicate firm-year observations based on the combination of the Worldscope Permanent ID (`item6105`) and the fiscal year (`year_`) and confirmed that there were no duplicates present, ensuring the accuracy of the data for further analysis.
 As the original study does not specify the net income measure, `item1651` (Net Income before Preferred Dividends) was used as the variable for net income.
@@
 Each country should have at least 300 firm-year observations for certain key accounting variables, including total assets, sales, net income, and operating income.
-Secondly, firms with adequate consecutive data must be identified.
-Each firm must have income statement and balance sheet information for at least three consecutive years, where I focus on the same key variables as in the previous step.
+In this step, no countries were eliminated as all countries met the requirement, which aligns with the paper's overview of countries.
+
+Additionally, firms with adequate consecutive data must be identified.
+Each firm must have income statement and balance sheet information for at least three consecutive years, with all key accounting variables mentioned above present.
 If a firm had at least three consecutive years of complete data at any point, all its data entries were retained in the final dataset.
+Therefore, only those countries and firms that meet these criteria for all specified variables are retained in the dataset.
 

### 2024-08-07 code/python/prepare_data.py

@@ -22,10 +22,16 @@ def main():
         log.warning(f"Found {dup_obs.shape[0]} duplicate firm-year observations. Removing duplicates.")
         wrds_data = wrds_data.drop_duplicates(subset=['item6105', 'year_'], keep='first')
 
-    # Filter countries with at least 300 firm-year observations for key accounting variables
-    filtered_countries_data = filter_countries(wrds_data)
+    # Filter countries with at least 300 firm-year observations for key accounting variables as in paper
+    filtered_countries_data, eliminated_countries = filter_countries(wrds_data)
 
-    # Filter firms with at least three consecutive years of income statement and balance sheet information
+    # Print eliminated countries
+    if eliminated_countries:
+        log.info(f"Countries eliminated after filtration: {', '.join(eliminated_countries)}")
+    else:
+        log.info("No countries were eliminated after filtration.")
+
+    # Filter firms with at least three consecutive years of income statement and balance sheet information as in paper
     filtered_firms_data = filter_firms(filtered_countries_data)
 
     # Save the filtered dataset

### 2024-08-07 code/python/prepare_data.py
more modifications

@@ -32,7 +32,19 @@ def main():
         log.info("No countries were eliminated after filtration.")
 
     # Filter firms with at least three consecutive years of income statement and balance sheet information as in paper
+    initial_firm_count = len(filtered_countries_data['item6105'].unique())
+    initial_obs_count = len(filtered_countries_data)
+    
     filtered_firms_data = filter_firms(filtered_countries_data)
+
+    final_firm_count = len(filtered_firms_data['item6105'].unique())
+    final_obs_count = len(filtered_firms_data)
+
+    firms_dropped = initial_firm_count - final_firm_count
+    obs_dropped = initial_obs_count - final_obs_count
+
+    log.info(f"Firms dropped after filtration: {firms_dropped}")
+    log.info(f"Firm-year observations dropped after filtration: {obs_dropped}")
 
     # Save the filtered dataset
     filtered_firms_data.to_csv(cfg['prepared_data_save_path'], index=False)

### 2024-08-07 doc/paper.qmd
more modifications
update

@@
 
-Additionally, firms with adequate consecutive data must be identified.
+In the second filtration step firms with adequate consecutive data must be identified.
 Each firm must have income statement and balance sheet information for at least three consecutive years, with all key accounting variables mentioned above present.
@@
 Therefore, only those countries and firms that meet these criteria for all specified variables are retained in the dataset.
+
+During the preparation step, 8,265 firms and 20,521 firm-year observations (all due to second filtration step) were dropped, resulting in a final dataset with 18,040 firms and 123,469 firm-year observations.
+The differences in the numbers between the prepared dataset and the figures mentioned in the paper (70,955 firm-year observations and 8,616 non-financial firms) could be due to the assumptions listed in Section 3.
 

### 2024-08-07 code/python/prepare_data.py

@@ -15,6 +15,10 @@ def main():
 
     # Load the pulled data
     wrds_data = pd.read_csv(cfg['worldscope_sample_save_path'])
+    initial_obs_count_pulled = len(wrds_data)
+    initial_firm_count_pulled = len(wrds_data['item6105'].unique())
+    log.info(f"Initial number of observations after pulling data: {initial_obs_count_pulled}")
+    log.info(f"Initial number of firms after pulling data: {initial_firm_count_pulled}")
 
     # Check for duplicate firm-year observations
     dup_obs = wrds_data[wrds_data.duplicated(subset=['item6105', 'year_'], keep=False)]
@@ -45,6 +49,9 @@ def main():
 
     log.info(f"Firms dropped after filtration: {firms_dropped}")
     log.info(f"Firm-year observations dropped after filtration: {obs_dropped}")
+
+    log.info(f"Number of observations after preparation: {final_obs_count}")
+    log.info(f"Number of firms after preparation: {final_firm_count}")
 
     # Save the filtered dataset
     filtered_firms_data.to_csv(cfg['prepared_data_save_path'], index=False)

### 2024-08-07 doc/paper.qmd
update paper

@@
 always_allow_html: yes
+number-sections: true 
 header-includes:
@@
 
-# Introduction
+# Introduction {#sec-introduction}
 
@@
 
-# Research Design Choices and Assumptions
+# Research Design Choices and Assumptions {#sec-research_design_assumptions}
 
@@
 
-By following the steps provided in Section 4 and adhering to the assumptions made, I successfully replicated the analysis and produced the required table.
+By following the steps provided in @sec-replication_steps and adhering to the assumptions made, I successfully replicated the analysis and produced the required table.
 A thorough step-by-step approach, with each step clearly documented, helped to understand and verify the outputs.
 
-# Replication Steps
+# Replication Steps {#sec-replication_steps}
 ## Step 1: Pulling the Data and Managing the Databases
@@
 During the preparation step, 8,265 firms and 20,521 firm-year observations (all due to second filtration step) were dropped, resulting in a final dataset with 18,040 firms and 123,469 firm-year observations.
-The differences in the numbers between the prepared dataset and the figures mentioned in the paper (70,955 firm-year observations and 8,616 non-financial firms) could be due to the assumptions listed in Section 3.
+The differences in the numbers between the prepared dataset and the figures mentioned in the paper (70,955 firm-year observations and 8,616 non-financial firms) could be due to the assumptions such as variations in initial datasets, data updates, and filtering criteria listed in @sec-research_design_assumptions.
+However, the original study might have included additional data cleaning steps not explicitly mentioned, such as handling outliers, specific industry exclusions, or other criteria, which could affect the final counts.
 

### 2024-08-08 doc/paper.qmd
table 1 replication

@@
 
+Now that it is clear that the number of observations for this project is significantly higher than that in @Leuz_2003 study, in order to illustrate the differences and compare the firm-year observations, Table 1 from @Leuz_2003 was partially replicated (only columns on countries and firm/year observations) to distinguish specific discrepancies that could arise for certain countries.
+The following table representes the (partially) replicated Table 1.
+
 ## Step 3: Analysis Implementation and Table Reproduction

### 2024-08-08 code/python/prepare_data.py

@@ -58,6 +58,18 @@ def main():
 
     log.info("Preparing data for analysis ... Done!")
 
+    # Generate summary table for firm-year observations per country
+    summary_table = filtered_firms_data.groupby('item6026').size().reset_index(name='# Firm-years')
+    summary_table.columns = ['Country', '# Firm-years']
+    
+    # Compute mean, median, min, and max
+    summary_stats = summary_table['# Firm-years'].describe()[['mean', '50%', 'min', 'max']]
+    summary_stats.index = ['Mean', 'Median', 'Min', 'Max']
+    
+    # Display the summary table and stats
+    print(summary_table)
+    print(summary_stats)
+
 def filter_countries(df):
     key_vars = ['item2999', 'item1001', 'item1250', 'item1651']  # Total Assets, Net Sales, Operating Income, Net Income
     grouped = df.groupby('item6026')

### 2024-08-08 code/python/prepare_data.py
table 1

@@ -65,10 +65,17 @@ def main():
     # Compute mean, median, min, and max
     summary_stats = summary_table['# Firm-years'].describe()[['mean', '50%', 'min', 'max']]
     summary_stats.index = ['Mean', 'Median', 'Min', 'Max']
+    summary_stats = summary_stats.reset_index()
+    summary_stats.columns = ['Country', '# Firm-years']
     
-    # Display the summary table and stats
-    print(summary_table)
-    print(summary_stats)
+    # Combine the summary table and stats
+    table_1 = pd.concat([summary_table, summary_stats], ignore_index=True)
+    
+    # Save Table 1 to a CSV file
+    table_1_save_path = 'data/generated/table_1.csv'
+    table_1.to_csv(table_1_save_path, index=False)
+    
+    print("Table 1 saved to", table_1_save_path)
 
 def filter_countries(df):
     key_vars = ['item2999', 'item1001', 'item1250', 'item1651']  # Total Assets, Net Sales, Operating Income, Net Income

### 2024-08-08 code/python/prepare_data.py
correct path saving

@@ -72,10 +72,9 @@ def main():
     table_1 = pd.concat([summary_table, summary_stats], ignore_index=True)
     
     # Save Table 1 to a CSV file
-    table_1_save_path = 'data/generated/table_1.csv'
-    table_1.to_csv(table_1_save_path, index=False)
+    table_1.to_csv(cfg['table_1_save_path'], index=False)
     
-    print("Table 1 saved to", table_1_save_path)
+    print("Table 1 saved to", cfg['table_1_save_path'])
 
 def filter_countries(df):
     key_vars = ['item2999', 'item1001', 'item1250', 'item1651']  # Total Assets, Net Sales, Operating Income, Net Income

### 2024-08-08 code/python/prepare_data.py
update preparation file

@@ -5,6 +5,7 @@
 # ------------------------------------------------------------------------------
 
 import pandas as pd
+import pickle
 from utils import read_config, setup_logging
 
 log = setup_logging()
@@ -71,10 +72,14 @@ def main():
     # Combine the summary table and stats
     table_1 = pd.concat([summary_table, summary_stats], ignore_index=True)
     
-    # Save Table 1 to a CSV file
-    table_1.to_csv(cfg['table_1_save_path'], index=False)
+    # Save Table 1 to a pickle file
+    results = {
+        "table_1": table_1
+    }
+    with open(cfg['table_1_save_path'], 'wb') as f:
+        pickle.dump(results, f)
     
-    print("Table 1 saved to", cfg['table_1_save_path'])
+    #print("Table 1 saved to", cfg['table_1_save_path'])
 
 def filter_countries(df):
     key_vars = ['item2999', 'item1001', 'item1250', 'item1651']  # Total Assets, Net Sales, Operating Income, Net Income

### 2024-08-08 doc/paper.qmd
table 1
pickle output path

@@
 
-## delete or take fikirs below
-```{python}
-#| echo: false
-#| output: false
-
-import pickle
-
-with open('../output/results.pickle', 'rb') as f:
-        results = pickle.load(f)
-
-min_fyear, max_fyear, no_unique_firms = results['Desc information'].values()
-
-def rename_var_to_label(table, italic=True):
-    for var_name, var_label in results['Variable Names'].items():
-        var_label = f'\\textit{{{var_label}}}' if italic else var_label
-        table = table.replace(var_name, var_label)
-    return table
-```
-
-## from fikir
 ```{python}
@@
 
-with open('../output/results.pickle', 'rb') as f:
+with open('data/generated/table_1.pickle', 'rb') as f:
     results = pickle.load(f)
@@
 
+```{python}
+#| echo: false
+#| output: asis
+
+print(prep_latex_table(
+    results['table_1'],
+    caption='Table 1: Number of Firm-Year Observations per Country',
+    label='tab:table_1'
+))
+```
+
 ## Step 3: Analysis Implementation and Table Reproduction

### 2024-08-08 doc/paper.qmd
updated documentation

@@
 
-with open('data/generated/table_1.pickle', 'rb') as f:
+with open('../data/generated/table_1.pickle', 'rb') as f:
     results = pickle.load(f)
 
-def escape_underscores(df):
-    df.columns = [col.replace('_', '\_') for col in df.columns]
+def escape_special_characters(df):
+    df.columns = [col.replace('_', '\\_').replace('#', '\\#') for col in df.columns]
     return df
@@
 def prep_latex_table(df, caption=None, label=None):
-    df = escape_underscores(df.reset_index())
+    df = escape_special_characters(df.reset_index())
     num_columns = len(df.columns)
@@
 The replicated table is shown below, followed by replication steps documentation, a detailed comparison, and discussion of the results.
-
-## delete from fikir.
-```{python}
-#| echo: false
-#| output: asis
-
-print(prep_latex_table(
-    results['replication_results'],
-    caption='Replicated table - P/B Ratios And Future Residual Income',
-    label='tab:replicated_penman'
-))
-```
 
@@
 
-## delete. cell 3 on Descriptive Statistics Table
-```{python}
-#| label: fig-boxplot_smp
-#| fig-cap: "Distribution of Discretionary Accruals over Time"
-#| echo: false
-#| fig-format: svg
-
-results['fig_boxplot_smp']
-```
-
-```{python}
-#| label: fig-scatter_dd_lnta
-#| fig-cap: "Dechow and Dichev DA and Firm Size"
-#| echo: false
-#| fig-format: svg
-
-results['fig_scatter_dd_lnta']
-```
-
-```{python}
-#| echo: false
-#| output: asis
-
-lat_tab = rename_var_to_label(results['Descriptive Statistics']).split('\n')
-lat_tab[4] = '\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]'
-lat_tab[6] = '\\hline\\\\[-1.8ex]'
-lat_tab[-4] = '\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]'
-
-print('\n'.join([
-    '\\begin{table}[!htbp]',  
-    '\caption{\label{tab:descriptives}Descriptive Statistics}',
-    '\label{tab:tab-descriptives-sample}',
-    '\\begin{adjustbox}{width=\\textwidth}',
-    '\\begin{threeparttable}',
-    '\n'.join(lat_tab[3:-2]),
-    '\\begin{tablenotes}[flushleft]',
-    '\setlength{\labelsep}{0pt}',
-    '\\footnotetext',
-    f'\\item Note: The data is obtained from the Compustat U.S. as provided by WRDS. The sample covers the period {int(min_fyear)} to {int(max_fyear)} and {no_unique_firms:,} unique firms.',
-    '\end{tablenotes}',
-    '\end{threeparttable}',
-    '\end{adjustbox}',
-    '\\end{table}',
-]))
-```
-
-```{python}
-#| echo: false
-#| output: asis
-
-lat_tab = rename_var_to_label(results['Correlations'], False).split('\n')
-
-lat_tab[2] = '\\\\[-0.9ex]\\hline \\hline \\\\[-0.9ex]'
-lat_tab[4] = '\\hline\\\\[-1.8ex]'
-lat_tab[12] = '\\\\[-0.9ex]\\hline \\hline \\\\[-0.9ex]'
-lat_tab[14] = '\\begin{tablenotes}[flushleft]'
-
-print('\n'.join([
-    "\\begin{table}[!htbp]",
-    "\centering",
-    "\caption{\label{tab:CorrTable}Correlations}",
-    '\label{tab:corr}',
-    '\n'.join(lat_tab),
-    "\end{table}"
-]))
-```
-
-```{python}
-#| echo: false
-#| output: asis
-
-print('\n'.join([
-    '\\begin{table}[!htbp] \\centering',
-    '\\caption{\\label{tab:reg}Regressions}', 
-    '\label{tab:reg}',
-    rename_var_to_label('\n'.join(results['Regression'].split('\n')[3:-1])),
-    "\end{table}"
-]))
-```
-
-# Results
-##Cross-check your results with the original table to ensure accuracy.
-Compare your results with the results in the paper.
-If you notice differences, discuss where they may originate.
-Make sure to be as specific as possible.
-This could involve looking at the data selection, calculation methods, or any assumptions that might differ from the original; changes in the data over time, different interpretations of the methodology, or limitations in the data sources used.
-Consider whether the differences are significant and what they might mean for the conclusions drawn in the paper.
-
-The replicated table shows some similarities with the original table by @Leuz_2003 in that . These findings are in line with valuation theory that .
-
-However, the patterns are less clear in the replicated table (in particular ), which stands in contrast to @Leuz_2003.
-
-# Conclusion
-
-##gpt: Summarize the key findings and their significance. take results and insert them here
-
-This project demonstrates the effectiveness of using an open science and collaborative workflow for analyzing earnings management measures across various countries.
-By following a step-by-step approach and using the TRR 266 Template for Reproducible Empirical Accounting Research, I was able to replicate an empirical table from @Leuz_2003 and provide insights into the trends of earnings management practices, which vary significantly across countries, similar to @Leuz_2003.
-
-The study by @Leuz_2003 documents systematic differences in the level of earnings management across 31 countries.
-It suggests that insiders use EM to conceal firm performance and protect their private control benefits and that stronger investor protection reduces earnings management by limiting insiders’ ability to gain private benefits.
-By conducting a descriptive cluster analysis after descriptive statistics that was replicated in this project, the paper identifies groupings of countries with similar institutional characteristics and demonstrates that earnings management varies systematically across these institutional clusters.
-
-The analysis revealed that
-
 The replication yields mixed results. While

### 2024-08-08 code/python/prepare_data.py

@@ -79,7 +79,7 @@ def main():
     with open(cfg['table_1_save_path'], 'wb') as f:
         pickle.dump(results, f)
     
-    #print("Table 1 saved to", cfg['table_1_save_path'])
+    log.info(f"Table 1 saved to {cfg['table_1_save_path']}")
 
 def filter_countries(df):
     key_vars = ['item2999', 'item1001', 'item1250', 'item1651']  # Total Assets, Net Sales, Operating Income, Net Income

### 2024-08-08 doc/paper.qmd
update

--- 2024-08-08 doc/paper.qmd
@@
 
-def prep_latex_table(df, caption=None, label=None):
-    df = escape_underscores(df.reset_index())
+def prep_latex_table(df, caption=None, label=None, fontsize='\\fontsize{11}{13}\\selectfont'):
+    df = escape_special_characters(df.reset_index())
     num_columns = len(df.columns)
@@
         "\\centering",
-        "\\resizebox{\\textwidth}{!}{%",
+        fontsize,  # Set font size for the table
         latex_table,
-        "}",
         f"\\caption{{{caption}}}" if caption else "",
@@
 Now that it is clear that the number of observations for this project is significantly higher than that in @Leuz_2003 study, in order to illustrate the differences and compare the firm-year observations, Table 1 from @Leuz_2003 was partially replicated (only columns on countries and firm/year observations) to distinguish specific discrepancies that could arise for certain countries.
-The following table representes the (partially) replicated Table 1.
+The following representes the (partially) replicated Table 1.
 
 ```{python}
+#| label: table_1 
 #| echo: false
@@
     results['table_1'],
-    caption='Table 1: Number of Firm-Year Observations per Country',
+    caption='Number of Firm-Year Observations per Country',
     label='tab:table_1'

### 2024-08-08 code/python/prepare_data.py

@@ -71,6 +71,9 @@ def main():
     
     # Combine the summary table and stats
     table_1 = pd.concat([summary_table, summary_stats], ignore_index=True)
+
+    # Remove decimal places
+    table_1['# Firm-years'] = table_1['# Firm-years'].apply(lambda x: int(round(x)))
     
     # Save Table 1 to a pickle file
     results = {

### 2024-08-08 doc/paper.qmd
step 2 finished

@@
 Now that it is clear that the number of observations for this project is significantly higher than that in @Leuz_2003 study, in order to illustrate the differences and compare the firm-year observations, Table 1 from @Leuz_2003 was partially replicated (only columns on countries and firm/year observations) to distinguish specific discrepancies that could arise for certain countries.
-The following representes the (partially) replicated Table 1.
+
+The results in Table 2 show a higher number of firm-year observations for almost all countries compared to the paper by @Leuz_2003, likely due to database updates leading to a more comprehensive dataset.
+The trend in the number of firm-year observations across countries is consistent with the paper, with larger capital markets like Japan, the United Kingdom, and the United States having the highest numbers.
+The distribution of observations in my dataset shows higher mean and median values compared to the paper by @Leuz_2003, indicating a greater overall number of firm-year observations, while the minimum values are relatively close, suggesting that the overall distribution of observations across countries remains broadly similar.
+
+However, there are significant discrepancies, particularly for the United States, which has 41,585 observations compared to 3,792 in the original study, possibly due to more comprehensive data or changes in dataset scope.
+@Leuz_2003 focuse only on S&P 500 companies for the U.S., while this analysis uses a more comprehensive dataset from the Worldscope database, which includes detailed financial statement data and profile data on public companies globally.
+This broader scope results in a higher number of firm-year observations for the United States.
+Despite the differences in absolute numbers, the relative trends across countries align with the paper.
 

### 2024-08-09 doc/paper.qmd
update paper.qmd

@@
 In doing so, it documents and discusses the research design choices made and the variations between the original and reproduced results.
-This code base, adapted from TREAT, should give you an overview on how the template is supposed to be used for my specific project and how to structure a reproducible empirical project.
+This code base, adapted from TREAT, should give you an overview on how the template is supposed to be used for this specific project and how to structure a reproducible empirical project.
   | \vspace{6cm}
@@
 
-The following paper illustrates how to use open science tools in empirical accounting research.
-
+The aim of the following paper is to illustrate how to use open science tools in empirical accounting research.
 The original study by @Leuz_2003 examines systematic differences in earnings management (EM) across 31 countries and proposes that these differences are influenced by the level of investor protection.
@@
 
-This paper presents details on a replication of Table 2 Panel A from the seminal paper by @Leuz_2003.
+This paper presents details on a replication of Table 2 Panel A from the seminal study by @Leuz_2003.
 The original table provides descriptive statistics for four individual earnings management measures and the aggregate earnings management score across various countries, highlighting significant differences in EM practices between Continental Europe, Asia, and Anglo-American countries.
@@
 
-The analysis includes pulling and filtering the relevant financial data, identifying relevant companies and countries based on the authors’ requirements, calculating the earnings management measures for each country, and examining the relationship between each country’s aggregate earnings management score and its ranking based on made explicit assumptions whenever the paper is not clear on how to proceed.
-
-## delete. from mine
-Note that intermediate check steps have been omitted for brevity.
-The full code, including intermediate print steps to verify outputs, is presented in `code/python/Assignment1_analyze_residual_income.py`.
+The analysis includes pulling and filtering the relevant financial data, identifying relevant companies and countries based on the authors’ requirements, calculating the EM measures for each country, and examining the relationship between each country’s aggregate earnings management score and its ranking based on explicit assumptions made whenever the paper by @Leuz_2003 is not clear on how to proceed.
+The complete computation code for EM measures is available in `code/python/do_analysis.py` for detailed review if necessary.
+Code snippets will be presented only in exceptional cases to highlight specific points.
+@
 
@@
     results['table_1'],
-    caption='Number of Firm-Year Observations per Country',
+    caption='Replicated table - Number of Firm-Year Observations per Country',
     label='tab:table_1'

### 2024-08-09 doc/paper.qmd
update paper.qmd

@@
 The aim of the following paper is to illustrate how to use open science tools in empirical accounting research.
+
 The original study by @Leuz_2003 examines systematic differences in earnings management (EM) across 31 countries and proposes that these differences are influenced by the level of investor protection.
@@
 Code snippets will be presented only in exceptional cases to highlight specific points.
-@
 
-The replicated table is shown below, followed by replication steps documentation, a detailed comparison, and discussion of the results.
+The replicated table is shown below, followed by the research design assumptions in @sec-research_design_assumptions, documentation of the replication steps in @sec-replication_steps, a detailed comparison in @sec-results, and concluding remarks in @sec-conclusion.
 
@@
 
-# Results
+# Results {#sec-results}
 ##Cross-check your results with the original table to ensure accuracy.
@@
 
-# Conclusion
+# Conclusion {#sec-conclusion}
 

### 2024-08-09 doc/paper.qmd
update citing referencing

@@
   | This project uses the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) to provide an infrastructure for open science-oriented empirical projects.
-Leveraging external Worldscope data sets on financial data, the repository showcases a reproducible workflow that integrates Python scripts for data analysis.
-The project’s output demonstrates a comprehensive application of skills to replicate and extend the findings from the seminal paper by Leuz, Nanda, and Wysocki (2003), particularly in providing descriptive statistics for the four individual earnings management measures as well as the aggregate earnings management score across various countries.
+Leveraging Worldscope data on financial information, this repository showcases a reproducible workflow that integrates Python scripts for data analysis.
+The project’s output demonstrates a comprehensive application of skills to replicate and extend the findings from the seminal paper by Leuz, Nanda, and Wysocki (2003), particularly in providing descriptive statistics for four individual earnings management measures, as well as an aggregate earnings management score across various countries.
 In doing so, it documents and discusses the research design choices made and the variations between the original and reproduced results.
-This code base, adapted from TREAT, should give you an overview on how the template is supposed to be used for this specific project and how to structure a reproducible empirical project.
+This code base, adapted from TREAT, provides an overview of how the template can be used for this specific project and how to structure a reproducible empirical research project.
   | \vspace{6cm}
@@
 
-The aim of the following paper is to illustrate how to use open science tools in empirical accounting research.
+The aim of this paper is to illustrate the use of open science tools in empirical accounting research.
 
 The original study by @Leuz_2003 examines systematic differences in earnings management (EM) across 31 countries and proposes that these differences are influenced by the level of investor protection.
-Specifically, they argue that stronger investor protection reduces earnings management by limiting insiders’ ability to acquire private control benefits.
+Specifically, the paper argues that stronger investor protection reduces earnings management by limiting insiders’ ability to acquire private control benefits.
 
-This paper presents details on a replication of Table 2 Panel A from the seminal study by @Leuz_2003.
+This paper presents the replication of Table 2 Panel A from the seminal study by @Leuz_2003.
 The original table provides descriptive statistics for four individual earnings management measures and the aggregate earnings management score across various countries, highlighting significant differences in EM practices between Continental Europe, Asia, and Anglo-American countries.
 
-The project uses data from the Worldscope Database through Wharton Research Data Services (WRDS), focusing on the fiscal years 1990 to 1999 and covering multiple countries.
+The project uses data from the Worldscope Database via Wharton Research Data Services (WRDS), focusing on fiscal years 1990 to 1999 and covering multiple countries.
 The replication aims to follow the research design by @Leuz_2003 as closely as possible with the available data.
@@
 
-The analysis includes pulling and filtering the relevant financial data, identifying relevant companies and countries based on the authors’ requirements, calculating the EM measures for each country, and examining the relationship between each country’s aggregate earnings management score and its ranking based on explicit assumptions made whenever the paper by @Leuz_2003 is not clear on how to proceed.
+The analysis includes pulling and filtering the relevant financial data, identifying relevant companies and countries based on the authors’ requirements, calculating the EM measures for each country, and examining the relationship between each country’s aggregate earnings management score and its ranking.
+Explicit assumptions were made whenever the paper by @Leuz_2003 was unclear on how to proceed.
 The complete computation code for EM measures is available in `code/python/do_analysis.py` for detailed review if necessary.
@@
 The replication is based on data pulled from the Worldscope database, specifically from the `wrds_ws_company` and `wrds_ws_funda` tables, which were merged for the analysis.
-The first table provides company profile information, including items such as ISIN, Worldscope Identifiers, company name, and the country where the company is domiciled @WRDS_WS_Company_2024.
-The latter table contains Fundamentals Annual data at the company-year level, including items such total assets, net income, and other relevant financial variables @WRDS_WS_Funda_2024.
+The first table provides company profile information, including items such as ISIN, Worldscope Identifiers, company name, and the country where the company is domiciled [@WRDS_WS_Company_2024].
+The latter table contains Fundamentals Annual data at the company-year level, including items such as total assets, net income, and other relevant financial variables [@WRDS_WS_Funda_2024].
 
@@
 
-1.
-The original paper references the November 2000 version of the Worldscope Database.
+1.	The original paper references the November 2000 version of the Worldscope Database.
 However, the data used for this analysis represents the latest available version, updated in July 2024, with quarterly frequency updates [@WRDS_2024].
 Due to potential adjustments and updates made to the database since November 2000, there may be differences between the databases that could affect the results.
-For example, companies may restate financials after the original reporting period, so that these restatements are reflected in the later database version rather than the historical one.
+For example, companies may restate financials after the original reporting period, so these restatements are reflected in the later database version rather than the historical one.
 Moreover, the data vendor Refinitiv regularly updates its databases to correct errors and add new information, which may be included in the later data but not in the November 2000 snapshot.
-2.
-The original paper outlines key terms that will be used in this project to ensure consistency and accuracy in the replication.
-@Leuz_2003 define earnings management as the manipulation of a firm's reported economic performance by insiders to deceive certain stakeholders or to affect contractual results.
-Authors describe investor protection as a key institutional factor that limits insiders' acquisition of private control benefits, thereby reducing their incentives to manage accounting earnings by ensuring strong and well-enforced rights for outside investors.
-Finally, private control benefits are the benefits that insiders can gain from controlling a firm, which can include financial gains or other advantages that are not shared with other stakeholders @Leuz_2003.
-3.
-While pulling the data for analysis, I encountered negative values for some key financial metrics such as operating income (`item1250`), or net income before preferred dividends (`item1651`).
+2.	The original paper outlines key terms that will be used in this project to ensure consistency and accuracy in the replication.
+@Leuz_2003 define earnings management as the manipulation of a firm’s reported economic performance by insiders to deceive certain stakeholders or to affect contractual outcomes.
+The authors describe investor protection as a key institutional factor that limits insiders’ acquisition of private control benefits, thereby reducing their incentives to manage accounting earnings by ensuring strong and well-enforced rights for outside investors.
+Finally, private control benefits are the advantages that insiders can gain from controlling a firm, which can include financial gains or other benefits not shared with other stakeholders [@Leuz_2003].
+3.	While pulling the data for analysis, I encountered negative values for some key financial metrics such as operating income (`item1250`) and net income before preferred dividends (`item1651`).
 The paper by @Leuz_2003 does not explicitly specify how to handle negative values in key financial metrics.
-For the purpose of this replication study, I will include negative values in the analysis.
+For the purpose of this replication study, I included negative values in the analysis.
 Including these values ensures that the analysis captures the full spectrum of earnings management activities across different countries.
-4.
-Another potential source of discrepancies between the original and replicated tables may be the choice of variables pulled from Worldscope.
-For example, in the `wrds_ws_funda table`, both `item1151` and `item4051` are named “DEPRECIATION, DEPLETION AND AMORTIZATION” @WRDS_WS_Funda_2024.
-I chose `item1151` from the Income Statement rather than `item4051` from the Cash Flow Statement, based on the Excel Industrials Template by @WRDS_Balancing_Model_2024.
+4.	Another potential source of discrepancies between the original and replicated tables may be the choice of variables pulled from Worldscope.
+For example, in the `wrds_ws_funda table`, both `item1151` and `item4051` are labeled as “DEPRECIATION, DEPLETION AND AMORTIZATION” [@WRDS_WS_Funda_2024].
+I chose `item1151` from the Income Statement rather than `item4051` from the Cash Flow Statement, based on the *Worldscope Balancing Model - Industrials* by Wharton Research Data Services [-@WRDS_Balancing_Model_2024].
 Since the authors do not specify the choice of variables used from the database, this could cause differences in the results.
-5.
-The EM measures are based on scaled variables (e.g., operating cash flow scaled by lagged total assets).
+5.	The EM measures are based on scaled variables (e.g., operating cash flow scaled by lagged total assets).
 As such, the currency of the relevant data items should not affect the results as long as the same currency is used consistently for both the numerator and the denominator.
@@
 Additionally, according to a document by [@Thomson_Financial_2007, p.20], all Worldscope data is consistently reported in the local currency of each firm’s country of domicile, eliminating the need for currency conversions in this project.
-333. ## delete. mine 
-333.
-It is assumed that Penman (2013) restricted the possible P/B values to a range of 0 to 7 to exclude outliers.
-Hence, extreme P/B values that are negative and very high (greater than 7) are excluded to focus the analysis on firms with more stable and reasonable valuations, reducing the impact of outliers.
 
@@
 # Replication Steps {#sec-replication_steps}
+
 ## Step 1: Pulling the Data and Managing the Databases
-In contrast to Assignment I, where the data was provided externally, Assignment III involves additionally pulling data directly from the Worldscope database, merging relevant tables, and preparing the data for further analysis from raw data to final output.
+Unlike Assignment I, where the data was provided externally, Assignment III involves pulling data directly from the Worldscope database, merging relevant tables, and preparing the data for further analysis from raw data to final output.
 
-To ensure the integrity of the data, I filtered out rows with empty `item6105` (Worldscope Permanent ID) values, as this identifier is critical for firm/year level filtering in the data preparation step.
-In total, 125 observations from the dynamic data and 10,306 observations from the static data were removed.
-To compile the dataset, the dynamic and static datasets were merged on the `item6105` identifier, representing the unique Worldscope Permanent ID.
-WRDS advises using this identifier consistently within Worldscope data since it remains stable over time [@WRDS_Overview_2024].
-An inner join was used for this merge, because this approach ensures that only the complete and consistent data from both tables is retained.
+To ensure data integrity, I filtered out rows with empty `item6105` (Worldscope Permanent ID) values, as this identifier is crucial for firm/year-level filtering in the data preparation step.
+A total of 125 observations from the dynamic data and 10,306 observations from the static data were removed.
+The dynamic and static datasets were then merged on the `item6105` identifier, representing the unique Worldscope Permanent ID.
+According to @WRDS_Overview_2024, this identifier should be used consistently within Worldscope data, as it remains stable over time.
+An inner join was used for this merge to ensure that only complete and consistent data from both tables were retained.
 
-Moreover, an additional filter is applied to retain only company rows.
-This is achieved by selecting rows where the `item6100` field equals 'C', indicating that the Worldscope Identifier represents a company.
-This step ensures that the analysis includes only company data, excluding averages, exchange rates, securities, or stock indices, as indicated by [@WRDS_Overview_2024].
+Additionally, a filter was applied to retain only company rows.
+This was achieved by selecting rows where the `item6100` field equals ‘C’, indicating that the Worldscope Identifier represents a company.
+This step ensures that the analysis includes only company data, excluding averages, exchange rates, securities, or stock indices, as advised by @WRDS_Overview_2024.
 
-As required by @Leuz_2003, financial institutions are removed from the analysis based on their SIC codes.
-This is done by filtering out rows where the `item7021` identifier, representing the SIC code, falls within the range of 6000 to 6999.
-Hence, the dataset focuses only on non-financial companies, aligning with the methodology of the original paper.
+In line with the methodology used by @Leuz_2003, financial institutions were removed from the analysis based on their SIC codes.
+This was done by filtering out rows where the `item7021` identifier, representing the SIC code, falls within the range of 6000 to 6999.
+This ensures that the dataset focuses solely on non-financial companies.
 
-Finally, additional filtering ensures that only data from the 31 countries, as given in the paper, is included. These filtering steps were applied to reduce the dataset size and improve the workflow.
+Further filtering ensured that only data from the 31 countries specified in the paper were included. These steps were applied to reduce the dataset size and improve the workflow.
 
-Notably, the configuration file utilizes additional refinement as proposed by [@WRDS_Overview_2024].
-Using `A` (Annual) in `freq` variable ensures that the data represents the financial information reported on an annual basis, which is consistent with the paper’s methodology.
+Moreover, the configuration file utilized additional refinements as proposed by @WRDS_Overview_2024.
+Specifically, using "A" (Annual) in the `freq` variable ensured that the data represents financial information reported on an annual basis, consistent with the methodology outlined in the paper by @Leuz_2003.
 This excludes data reported on a current, daily, or quarterly basis.
 
-After retaining only relevant company data, filtering out financial institutions, and focusing on specified countries, the processed data was then saved to a CSV file at the path `data/pulled` specified in the configuration file `config/pull_data_cfg​`.
+After retaining only relevant company data, filtering out financial institutions, and focusing on the specified countries, the processed data was saved to a CSV file in the `data/pulled` directory, as specified in the configuration file `config/pull_data_cfg`​.
 
 ## Step 2: Data Preparation
-In order to verify the pulle data, I checked the dataset for duplicate firm-year observations based on the combination of the Worldscope Permanent ID (`item6105`) and the fiscal year (`year_`) and confirmed that there were no duplicates present, ensuring the accuracy of the data for further analysis.
-As the original study does not specify the net income measure, `item1651` (Net Income before Preferred Dividends) was used as the variable for net income.
-This choice aligns with the final net income figure reported in the income statement, based on the Excel Industrials Template by @WRDS_Balancing_Model_2024.
+To verify the pulled data, I checked the dataset for duplicate firm-year observations based on the combination of the Worldscope Permanent ID (`item6105`) and the fiscal year (`year_`).
+I confirmed that no duplicates were present, ensuring the accuracy of the data for further analysis.
+As the original study does not specify the net income measure, I used `item1651` (Net Income before Preferred Dividends) as the variable for net income.
+This choice aligns with the final net income figure reported in the income statement, based on *Worldscope Balancing Model - Industrials* by Wharton Research Data Services [-@WRDS_Balancing_Model_2024].
 
-To prepare the data sample in line with the methodology outlined in the paper, it is essential to follow the requirements, as instructed by @Leuz_2003.
-Firstly, countries with sufficient firm-year observations must be filtered.
-Each country should have at least 300 firm-year observations for certain key accounting variables, including total assets, sales, net income, and operating income.
-In this step, no countries were eliminated as all countries met the requirement, which aligns with the paper's overview of countries.
+To prepare the data sample in line with the methodology outlined in the paper, it was essential to follow the requirements set by @Leuz_2003.
+First, countries with sufficient firm-year observations were filtered.
+Each country needed at least 300 firm-year observations for key accounting variables, including total assets, sales, net income, and operating income.
+In this step, no countries were eliminated, as all countries met the requirement, which aligns with the paper’s overview of countries (31 countries in total).
 
-In the second filtration step firms with adequate consecutive data must be identified.
-Each firm must have income statement and balance sheet information for at least three consecutive years, with all key accounting variables mentioned above present.
+In the second filtration step, firms with adequate consecutive data were identified.
+Each firm had to have income statement and balance sheet information for at least three consecutive years, with all key accounting variables mentioned above present.
 If a firm had at least three consecutive years of complete data at any point, all its data entries were retained in the final dataset.
-Therefore, only those countries and firms that meet these criteria for all specified variables are retained in the dataset.
+Therefore, only those countries and firms that met these criteria for all specified variables were retained in the dataset.
 
-During the preparation step, 8,265 firms and 20,521 firm-year observations (all due to second filtration step) were dropped, resulting in a final dataset with 18,040 firms and 123,469 firm-year observations.
-The differences in the numbers between the prepared dataset and the figures mentioned in the paper (70,955 firm-year observations and 8,616 non-financial firms) could be due to the assumptions such as variations in initial datasets, data updates, and filtering criteria listed in @sec-research_design_assumptions.
-However, the original study might have included additional data cleaning steps not explicitly mentioned, such as handling outliers, specific industry exclusions, or other criteria, which could affect the final counts.
+During the preparation step, 8,265 firms and 20,521 firm-year observations (all due to the second filtration step) were dropped, resulting in a final dataset with 18,040 non-financial firms and 123,469 firm-year observations.
+The differences in the numbers between the prepared dataset and the figures mentioned in the paper (70,955 firm-year observations and 8,616 non-financial firms) could be due to assumptions outlined in @sec-research_design_assumptions, such as variations in initial datasets, data updates, and filtering criteria.
+However, the original paper might have included additional data cleaning steps not explicitly mentioned, such as handling outliers, specific industry exclusions, or other criteria, which could affect the final counts.
 
-Now that it is clear that the number of observations for this project is significantly higher than that in @Leuz_2003 study, in order to illustrate the differences and compare the firm-year observations, Table 1 from @Leuz_2003 was partially replicated (only columns on countries and firm/year observations) to distinguish specific discrepancies that could arise for certain countries.
-
-The results in Table 2 show a higher number of firm-year observations for almost all countries compared to the paper by @Leuz_2003, likely due to database updates leading to a more comprehensive dataset.
-The trend in the number of firm-year observations across countries is consistent with the paper, with larger capital markets like Japan, the United Kingdom, and the United States having the highest numbers.
+Given that the number of observations in this project is significantly higher than in the original study, Table 1 from @Leuz_2003 was partially replicated (focusing only on columns related to countries and firm-year observations) to highlight specific discrepancies that could arise for certain countries.
+The results in the table below show a higher number of firm-year observations for almost all countries compared to the paper by @Leuz_2003, likely due to database updates leading to a more comprehensive dataset.
+The trend in the number of firm-year observations across countries is consistent with the original study, with larger capital markets like Japan, the United Kingdom, and the United States having the highest numbers.
 The distribution of observations in my dataset shows higher mean and median values compared to the paper by @Leuz_2003, indicating a greater overall number of firm-year observations, while the minimum values are relatively close, suggesting that the overall distribution of observations across countries remains broadly similar.
@@
 However, there are significant discrepancies, particularly for the United States, which has 41,585 observations compared to 3,792 in the original study, possibly due to more comprehensive data or changes in dataset scope.
-@Leuz_2003 focuse only on S&P 500 companies for the U.S., while this analysis uses a more comprehensive dataset from the Worldscope database, which includes detailed financial statement data and profile data on public companies globally.
+@Leuz_2003 focused only on S&P 500 companies for the U.S., while this analysis uses a more comprehensive dataset from the Worldscope database, which includes detailed financial statement data and profile data on public companies globally.
 This broader scope results in a higher number of firm-year observations for the United States.
@@
 # Results {#sec-results}
+
 ##Cross-check your results with the original table to ensure accuracy.

### 2024-08-09 doc/paper.qmd
step 3

@@
 However, the original paper might have included additional data cleaning steps not explicitly mentioned, such as handling outliers, specific industry exclusions, or other criteria, which could affect the final counts.
-
+## table 2 correctly referenced?
 Given that the number of observations in this project is significantly higher than in the original study, Table 1 from @Leuz_2003 was partially replicated (focusing only on columns related to countries and firm-year observations) to highlight specific discrepancies that could arise for certain countries.
-The results in the table below show a higher number of firm-year observations for almost all countries compared to the paper by @Leuz_2003, likely due to database updates leading to a more comprehensive dataset.
+The results in the table 2 below show a higher number of firm-year observations for almost all countries compared to the paper by @Leuz_2003, likely due to database updates leading to a more comprehensive dataset.
 The trend in the number of firm-year observations across countries is consistent with the original study, with larger capital markets like Japan, the United Kingdom, and the United States having the highest numbers.
@@
 ## Step 3: Analysis Implementation and Table Reproduction
-In this replication step, implement the calculations for the EM measures as described in the methodology of the paper, ensuring that all statistical methods and groupings are accurately replicated.
+In this replication step, the calculation of the EM measures is implemented according to the methodology described by @Leuz_2003.
+The study employs four distinct proxies that capture different dimensions of earnings management, such as earnings smoothing and accrual manipulations.
+These measures are used to assess the pervasiveness of earnings management across countries.
+This step ensures that all statistical methods and groupings are accurately replicated, following the paper’s approach to capture the extent to which corporate insiders use accounting discretion to obscure their firm’s true economic performance.
+
+### Smoothing Reported Operating Earnings Using Accruals (EM1)
+`EM1` is a measure that assesses how much insiders (e.g., company executives) smooth or stabilize reported earnings by manipulating accruals.
+The underlying idea is that firms may attempt to reduce fluctuations in their reported operating income to present a more stable financial performance to investors.
+This measure indicates the degree to which insiders exercise accounting discretion to smooth reported earnings, with lower values of EM1 suggesting more aggressive earnings smoothing.
+
+To calculate this measure:
+1.	Accruals are computed using changes in various components of a firm’s balance sheet and income statement, such as total current assets, cash, current liabilities, short-term debt, taxes payable, and depreciation.
+$$
+\text{Accruals}_{it} = (\Delta \text{CA}_{it} - \Delta \text{Cash}_{it}) - (\Delta \text{CL}_{it} - \Delta \text{STD}_{it} - \Delta \text{TP}_{it}) - \text{Dep}_{it}
+$$
+
+$$
+\begin{aligned}
+where:  \\
+\Delta \text{CA}_{it} & = \text{Change in total current assets} \\
+\Delta \text{Cash}_{it} & = \text{Change in cash and cash equivalents} \\
+\Delta \text{CL}_{it} & = \text{Change in total current liabilities} \\
+\Delta \text{STD}_{it} & = \text{Change in short-term debt included in current liabilities} \\
+\Delta \text{TP}_{it} & = \text{Change in income taxes payable} \\
+\text{Dep}_{it} & = \text{Depreciation and amortization expense}
+\end{aligned}
+$$
+
+2.	The operating cash flow is indirectly derived from operating income minus accruals
+$$
+\text{CFO}_{it} = \text{Operating Income}_{it} - \text{Accruals}_{it}
+$$
+
+3.	EM1 is then calculated as the country’s median ratio of the firm-level standard deviations of operating income and operating cash flow.
+Both these standard deviations are scaled by lagged total assets to control for differences in firm size.
+$$
+\text{EM1} = \text{Median}\left(\frac{\sigma(\text{Operating Income}_{it})}{\sigma(\text{Operating Cash Flow}_{it})}\right) \quad \text{scaled by lagged total assets}
+$$
+
+### Smoothing and the Correlation Between Changes in Accounting Accruals and Operating Cash Flows (EM2)
+
+### Discretion in Reported Earnings: The Magnitude of Accruals (EM3)
+
+### Discretion in Reported Earnings: Small Loss Avoidance (EM4)
+
+### Aggregate Measure of Earnings Management
 

### 2024-08-09 doc/paper.qmd
do_analysis.py code

@@
 
+**CFO**: Operating Cash Flow  
 **EM**: Earnings Management  
@@
 `EM1` is a measure that assesses how much insiders (e.g., company executives) smooth or stabilize reported earnings by manipulating accruals.
-The underlying idea is that firms may attempt to reduce fluctuations in their reported operating income to present a more stable financial performance to investors.
-This measure indicates the degree to which insiders exercise accounting discretion to smooth reported earnings, with lower values of EM1 suggesting more aggressive earnings smoothing.
+The concept is that firms may attempt to reduce fluctuations in their reported operating income to present a more stable financial performance to investors.
+This measure reflects the degree to which insiders use accounting discretion to smooth reported earnings, with lower values of EM1 suggesting more aggressive earnings smoothing.
 
 To calculate this measure:
-1.	Accruals are computed using changes in various components of a firm’s balance sheet and income statement, such as total current assets, cash, current liabilities, short-term debt, taxes payable, and depreciation.
-$$
-\text{Accruals}_{it} = (\Delta \text{CA}_{it} - \Delta \text{Cash}_{it}) - (\Delta \text{CL}_{it} - \Delta \text{STD}_{it} - \Delta \text{TP}_{it}) - \text{Dep}_{it}
-$$
+1.	Compute Accruals by using changes in various components of a firm’s balance sheet and income statement, such as total current assets, cash, current liabilities, short-term debt, taxes payable, and depreciation:
 
@@
 \begin{aligned}
-where:  \\
+\text{Accruals}_{it} &= (\Delta \text{CA}_{it} - \Delta \text{Cash}_{it}) - (\Delta \text{CL}_{it} - \Delta \text{STD}_{it} - \Delta \text{TP}_{it}) - \text{Dep}_{it} 
+& \\
+\text{where:} \\
 \Delta \text{CA}_{it} & = \text{Change in total current assets} \\
@@
 
-2.	The operating cash flow is indirectly derived from operating income minus accruals
+2.	Derive Operating Cash Flow (CFO) that is indirectly derived from operating income minus accruals:
+
 $$
-\text{CFO}_{it} = \text{Operating Income}_{it} - \text{Accruals}_{it}
+\text{CFO}{it} = \text{Operating Income}_{it} - \text{Accruals}_{it}
 $$
 
-3.	EM1 is then calculated as the country’s median ratio of the firm-level standard deviations of operating income and operating cash flow.
-Both these standard deviations are scaled by lagged total assets to control for differences in firm size.
+3.	Calculate EM1, which is determined as the country’s median ratio of the firm-level standard deviations of operating income and operating cash flow.
+Both standard deviations are scaled by lagged total assets to control for differences in firm size:
 $$

### 2024-08-09 code/python/do_analysis.py

@@ -1,133 +1,63 @@
 import pickle
 import pandas as pd
-import numpy as np
-from plotnine import ggplot, geom_boxplot, labs, aes, geom_bin2d, after_stat
-from panel_eda_helper_funcs import (prepare_correlation_table,
-                                    prepare_descriptive_table, PrepareRegressionTable,
-                                    escape_for_latex)
-from prepare_data import treat_outliers
-from theme_trr import theme_trr, scale_color_trr266_d, scale_fill_trr266_c
-
-from utils import setup_logging, read_config
-
-log = setup_logging()
-
 
 def main():
-    log.info('Performing main analysis...')
-    cfg = read_config('config/do_analysis_cfg.yaml')
-    smp = pd.read_csv(cfg['acc_sample'], dtype={'gvkey': str})
+    # Load prepared data
+    financial_data = load_data()
 
-    fig_boxplot_full = make_boxplot(smp)
+    # Calculate accruals
+    financial_data = calculate_accruals(financial_data)
 
-    smp_da = prep_smp_da(smp)
+    # Calculate Operating Cash Flow (CFO)
+    financial_data = calculate_cfo(financial_data)
 
-    fig_boxplot_smp = make_boxplot(smp_da)
+    # Calculate EM1
+    em1_results = calculate_em1(financial_data)
 
-    fig_scatter_md_dd = make_scatter_plot(
-        smp_da, 'mj_da', 'dd_da', 'Modified Jones DA', 'Dechow and Dichev DA'
-    )
+    # Save and print results
+    save_results(em1_results)
 
-    fig_scatter_dd_lnta = make_scatter_plot(
-        smp_da, 'ln_ta', 'dd_da', 'ln(Total Assets)', 'Dechow and Dichev DA'
-    )
+def load_data():
+    """
+    Load the prepared financial data from the previous step.
+    """
+    return pd.read_csv('data/generated/financial_data_prepared.csv')
 
-    fig_scatter_dd_roa = make_scatter_plot(
-        smp_da, 'ebit_avgta', 'dd_da', 'Return on Assets', 'Dechow and Dichev DA'
-    )
-    fig_scatter_dd_salesgr = make_scatter_plot(
-        smp_da, 'sales_growth', 'dd_da', 'Sales Growth', 'Dechow and Dichev DA'
-    )
+def calculate_accruals(df):
+    """
+    Calculate accruals using the provided formula.
+    """
[…]

### 2024-08-09 doc/paper.qmd
do analysis

@@
 The concept is that firms may attempt to reduce fluctuations in their reported operating income to present a more stable financial performance to investors.
-This measure reflects the degree to which insiders use accounting discretion to smooth reported earnings, with lower values of EM1 suggesting more aggressive earnings smoothing.
+This measure reflects the degree to which insiders use accounting discretion to smooth reported earnings, with lower values of EM1 suggesting more extensive earnings smoothing.
 
@@
 \Delta \text{TP}_{it} & = \text{Change in income taxes payable} \\
-\text{Dep}_{it} & = \text{Depreciation and amortization expense}
+\text{Dep}_{it} & = \text{Depreciation and amortization expense for firm i in year t} 
 \end{aligned}

### 2024-08-09 code/python/do_analysis.py

@@ -1,3 +1,4 @@
+# We start by loading the libraries that we will use in this analysis.
 import pickle
 import pandas as pd
 
@@ -21,13 +22,24 @@ def load_data():
     """
     Load the prepared financial data from the previous step.
     """
-    return pd.read_csv('data/generated/financial_data_prepared.csv')
+    df = pd.read_csv('data/generated/financial_data_prepared.csv')
+    print(df.head())  # Print the first 5 rows of the dataframe to show the header and a few entries
+    return df
 
 def calculate_accruals(df):
     """
     Calculate accruals using the provided formula.
+    If a firm does not report information on taxes payable or short-term debt,
+    then the change in both variables is assumed to be zero.
     """
-    df['Accruals'] = (df['item2201'].diff() - df['item2003'].diff()) - (df['item3101'].diff() - df['item3051'].diff() - df['item3063'].diff()) - df['item1151']
+    df['delta_CA'] = df['item2201'].diff()
+    df['delta_Cash'] = df['item2003'].diff()
+    df['delta_CL'] = df['item3101'].diff()
+    df['delta_STD'] = df['item3051'].diff().fillna(0)  # Apply fillna only for STD
+    df['delta_TP'] = df['item3063'].diff().fillna(0)   # Apply fillna only for TP
+    df['Dep'] = df['item1151']  # No fillna applied to Dep
+
+    df['Accruals'] = (df['delta_CA'] - df['delta_Cash']) - (df['delta_CL'] - df['delta_STD'] - df['delta_TP']) - df['Dep']
     return df
 
 def calculate_cfo(df):
@@ -47,6 +59,19 @@ def calculate_em1(df):
     df['EM1'] = df.groupby('item6105')['EM1'].transform(lambda x: x / df['item2999'].shift(1))  # Scale by lagged total assets
 
     country_em1 = df.groupby('item6026')['EM1'].median().reset_index()
+
+    # Round the EM1 results to three decimal places
+    country_em1['EM1'] = country_em1['EM1'].round(3)
+
+    # Calculate mean, median, standard deviation, min, max
+    summary_stats = country_em1['EM1'].agg(['mean', 'median', 'std', 'min', 'max']).round(3)
+
+    # Add the summary stats to the results
+    summary_stats = summary_stats.reset_index()
+    summary_stats.columns = ['Statistic', 'EM1']
+
+    country_em1 = pd.concat([country_em1, summary_stats], ignore_index=True)
+
     return country_em1
 
 def save_results(results):

### 2024-08-09 doc/paper.qmd
do analysis

@@
 
+During the accruals calculation, `NaN` values arose primarily due to the lack of prior year data needed to compute changes in key financial variables (e.g., total current assets, cash, current liabilities).
+This issue is most common in the first year of each firm’s data, where no prior year data exists.
+Furthermore, missing data in intermediate years (e.g., a firm has data for 1990, 1992, and 1993 but not 1991) can also result in `NaN` values for subsequent years due to the lack of necessary prior year information for the accruals calculation.
+
+In total, 53,231 rows with `NaN` values were identified and subsequently dropped from the dataset.
+@Leuz_2003 specify that for taxes payable (`delta_TP`) and short-term debt (`delta_STD`), missing data should be treated as zero, but for other variables, dropping these rows is consistent with financial practices to ensure accuracy.
+
 2.	Derive Operating Cash Flow (CFO) that is indirectly derived from operating income minus accruals:
@@
 $$
-\text{CFO}{it} = \text{Operating Income}_{it} - \text{Accruals}_{it}
+\text{CFO}_{it} = \text{Operating Income}_{it} - \text{Accruals}_{it}
 $$

### 2024-08-09 code/python/do_analysis.py

@@ -23,7 +23,6 @@ def load_data():
     Load the prepared financial data from the previous step.
     """
     df = pd.read_csv('data/generated/financial_data_prepared.csv')
-    print(df.head())  # Print the first 5 rows of the dataframe to show the header and a few entries
     return df
 
 def calculate_accruals(df):
@@ -32,14 +31,34 @@ def calculate_accruals(df):
     If a firm does not report information on taxes payable or short-term debt,
     then the change in both variables is assumed to be zero.
     """
-    df['delta_CA'] = df['item2201'].diff()
-    df['delta_Cash'] = df['item2003'].diff()
-    df['delta_CL'] = df['item3101'].diff()
-    df['delta_STD'] = df['item3051'].diff().fillna(0)  # Apply fillna only for STD
-    df['delta_TP'] = df['item3063'].diff().fillna(0)   # Apply fillna only for TP
-    df['Dep'] = df['item1151']  # No fillna applied to Dep
+
+    # Group by unique firm (item6105) and calculate diffs within each firm
+    df['delta_CA'] = df.groupby('item6105')['item2201'].diff()  # Change in total current assets
+    df['delta_Cash'] = df.groupby('item6105')['item2003'].diff()  # Change in cash and cash equivalents
+    df['delta_CL'] = df.groupby('item6105')['item3101'].diff()  # Change in total current liabilities
+    df['delta_STD'] = df.groupby('item6105')['item3051'].diff().fillna(0)  # Apply fillna only for STD
+    df['delta_TP'] = df.groupby('item6105')['item3063'].diff().fillna(0)  # Apply fillna only for TP
+    df['Dep'] = df['item1151']  # No fillna applied to Depreciation and amortization expense
 
     df['Accruals'] = (df['delta_CA'] - df['delta_Cash']) - (df['delta_CL'] - df['delta_STD'] - df['delta_TP']) - df['Dep']
+
+    # Identify rows with NaN in Accruals
+    nan_accruals = df[df['Accruals'].isna()]
+
+    # Print the first 10 rows with NaN in Accruals
+    print("First 10 rows with NaN in Accruals:")
+    print(nan_accruals.head(10))
+
+    # Count and print the number of NaN rows in Accruals
+    nan_count = df['Accruals'].isna().sum()
+    print(f"Number of rows with NaN in Accruals: {nan_count}")
+
+    # Drop rows where Accruals is NaN
[…]

### 2024-08-09 code/python/do_analysis.py
do analysis

@@ -68,30 +68,39 @@ def calculate_cfo(df):
     df['CFO'] = df['item1250'] - df['Accruals']
 
     print(df[['item6105', 'year_', 'Accruals', 'CFO']].head(10))
-    
+
     return df
 
 def calculate_em1(df):
     """
     Calculate EM1 for each firm and then take the country-level median.
     """
+    # Calculate the standard deviation of operating income for each firm (grouped by item6105)
     df['std_operating_income'] = df.groupby('item6105')['item1250'].transform('std')
+    
+    # Calculate the standard deviation of CFO for each firm (grouped by item6105)
     df['std_cfo'] = df.groupby('item6105')['CFO'].transform('std')
+    
+    # Calculate EM1 by dividing the standard deviation of operating income by the standard deviation of CFO
     df['EM1'] = df['std_operating_income'] / df['std_cfo']
-    df['EM1'] = df.groupby('item6105')['EM1'].transform(lambda x: x / df['item2999'].shift(1))  # Scale by lagged total assets
-
+    
+    # Scale EM1 by lagged total assets, grouped by firm (item6105) and shifting by one period
+    df['EM1'] = df.groupby('item6105')['EM1'].transform(lambda x: x / df['item2999'].shift(1))
+    
+    # Group by country (item6026) and take the median of EM1 for each country
     country_em1 = df.groupby('item6026')['EM1'].median().reset_index()
 
     # Round the EM1 results to three decimal places
     country_em1['EM1'] = country_em1['EM1'].round(3)
 
-    # Calculate mean, median, standard deviation, min, max
+    # Calculate mean, median, standard deviation, min, max of EM1 across countries
     summary_stats = country_em1['EM1'].agg(['mean', 'median', 'std', 'min', 'max']).round(3)
 
     # Add the summary stats to the results
     summary_stats = summary_stats.reset_index()
     summary_stats.columns = ['Statistic', 'EM1']
 
+    # Combine the country-level results and summary statistics
     country_em1 = pd.concat([country_em1, summary_stats], ignore_index=True)
 
     return country_em1

### 2024-08-09 doc/paper.qmd
do analysis more

@@
 This issue is most common in the first year of each firm’s data, where no prior year data exists.
-Furthermore, missing data in intermediate years (e.g., a firm has data for 1990, 1992, and 1993 but not 1991) can also result in `NaN` values for subsequent years due to the lack of necessary prior year information for the accruals calculation.
+Furthermore, missing data in intermediate years (e.g., a firm has data on cash for 1990, 1992, and 1993 but not 1991) can also result in `NaN` values for subsequent years due to the lack of necessary prior year information for the accruals calculation.
 
-In total, 53,231 rows with `NaN` values were identified and subsequently dropped from the dataset.
+In total, 53,231 rows with `NaN` values were identified ##and subsequently dropped from the dataset.
 @Leuz_2003 specify that for taxes payable (`delta_TP`) and short-term debt (`delta_STD`), missing data should be treated as zero, but for other variables, dropping these rows is consistent with financial practices to ensure accuracy.

### 2024-08-09 code/python/do_analysis.py

@@ -54,7 +54,7 @@ def calculate_accruals(df):
     print(f"Number of rows with NaN in Accruals: {nan_count}")
 
     # Drop rows where Accruals is NaN
-    df = df.dropna(subset=['Accruals'])
+    # df = df.dropna(subset=['Accruals'])
 
     # Print the first few rows to check the calculated accruals
     print(df[['item6105', 'year_', 'Accruals']].head(10))
@@ -64,9 +64,12 @@ def calculate_accruals(df):
 def calculate_cfo(df):
     """
     Calculate Operating Cash Flow (CFO) by subtracting accruals from operating income.
+    Handle NaN values where necessary.
     """
+    # We only calculate CFO where accruals are not NaN
     df['CFO'] = df['item1250'] - df['Accruals']
-
+    
+    # Print the first few rows to verify CFO calculations
     print(df[['item6105', 'year_', 'Accruals', 'CFO']].head(10))
 
     return df
@@ -77,15 +80,18 @@ def calculate_em1(df):
     """
     # Calculate the standard deviation of operating income for each firm (grouped by item6105)
     df['std_operating_income'] = df.groupby('item6105')['item1250'].transform('std')
-    
     # Calculate the standard deviation of CFO for each firm (grouped by item6105)
     df['std_cfo'] = df.groupby('item6105')['CFO'].transform('std')
     
-    # Calculate EM1 by dividing the standard deviation of operating income by the standard deviation of CFO
-    df['EM1'] = df['std_operating_income'] / df['std_cfo']
+    # For scaling, identify the first available year for each firm and get the total assets for the prior year
+    df['lagged_total_assets'] = df.groupby('item6105')['item2999'].shift(1)
     
-    # Scale EM1 by lagged total assets, grouped by firm (item6105) and shifting by one period
-    df['EM1'] = df.groupby('item6105')['EM1'].transform(lambda x: x / df['item2999'].shift(1))
+    # Scale the standard deviations individually by lagged total assets
+    df['scaled_std_operating_income'] = df['std_operating_income'] / df['lagged_total_assets']
+    df['scaled_std_cfo'] = df['std_cfo'] / df['lagged_total_assets']
+    
+    # Calculate EM1 as the ratio of the scaled standard deviations
+    df['EM1'] = df['scaled_std_operating_income'] / df['scaled_std_cfo']
     
     # Group by country (item6026) and take the median of EM1 for each country
     country_em1 = df.groupby('item6026')['EM1'].median().reset_index()

### 2024-08-09 doc/paper.qmd
EM1

@@
 $$
-\text{EM1} = \text{Median}\left(\frac{\sigma(\text{Operating Income}_{it})}{\sigma(\text{Operating Cash Flow}_{it})}\right) \quad \text{scaled by lagged total assets}
+\text{EM1} = \text{Median}\left(\frac{\sigma(\text{Operating Income}_{it})}{\sigma(\text{Operating Cash Flow}_{it})}\right) \quad \text{both separately scaled by lagged total assets}
 $$

### 2024-08-09 code/python/do_analysis.py

@@ -53,9 +53,6 @@ def calculate_accruals(df):
     nan_count = df['Accruals'].isna().sum()
     print(f"Number of rows with NaN in Accruals: {nan_count}")
 
-    # Drop rows where Accruals is NaN
-    # df = df.dropna(subset=['Accruals'])
-
     # Print the first few rows to check the calculated accruals
     print(df[['item6105', 'year_', 'Accruals']].head(10))
     
@@ -82,20 +79,27 @@ def calculate_em1(df):
     df['std_operating_income'] = df.groupby('item6105')['item1250'].transform('std')
     # Calculate the standard deviation of CFO for each firm (grouped by item6105)
     df['std_cfo'] = df.groupby('item6105')['CFO'].transform('std')
+    print("Standard Deviation of Operating Income and CFO (first 10 rows):")
+    print(df[['item6105', 'year_', 'std_operating_income', 'std_cfo']].head(10))
     
-    # For scaling, identify the first available year for each firm and get the total assets for the prior year
+    # For scaling, identify the total assets from the prior year (making use of year 0 as well. Otherwise would have dropped it in accruals step)
     df['lagged_total_assets'] = df.groupby('item6105')['item2999'].shift(1)
+    print("Lagged Total Assets (first 10 rows):")
+    print(df[['item6105', 'year_', 'item2999', 'lagged_total_assets']].head(10))
     
     # Scale the standard deviations individually by lagged total assets
     df['scaled_std_operating_income'] = df['std_operating_income'] / df['lagged_total_assets']
     df['scaled_std_cfo'] = df['std_cfo'] / df['lagged_total_assets']
+    print("Scaled Standard Deviation of Operating Income and CFO (first 15 rows):")
+    print(df[['item6105', 'year_', 'scaled_std_operating_income', 'scaled_std_cfo']].head(15))
     
     # Calculate EM1 as the ratio of the scaled standard deviations
     df['EM1'] = df['scaled_std_operating_income'] / df['scaled_std_cfo']
+    print("EM1 (first 15 rows):")
+    print(df[['item6105', 'year_', 'EM1']].head(15))
     
     # Group by country (item6026) and take the median of EM1 for each country
     country_em1 = df.groupby('item6026')['EM1'].median().reset_index()
-
     # Round the EM1 results to three decimal places
     country_em1['EM1'] = country_em1['EM1'].round(3)
 

### 2024-08-09 doc/paper.qmd
EM1 final

--- 2024-08-09 doc/paper.qmd
@@
 To calculate this measure:
+
 1.	Compute Accruals by using changes in various components of a firm’s balance sheet and income statement, such as total current assets, cash, current liabilities, short-term debt, taxes payable, and depreciation:
-
 $$
@@
 
-During the accruals calculation, `NaN` values arose primarily due to the lack of prior year data needed to compute changes in key financial variables (e.g., total current assets, cash, current liabilities).
-This issue is most common in the first year of each firm’s data, where no prior year data exists.
-Furthermore, missing data in intermediate years (e.g., a firm has data for 1990, 1992, and 1993 but not 1991) can also result in `NaN` values for subsequent years due to the lack of necessary prior year information for the accruals calculation.
-
-In total, 53,231 rows with `NaN` values were identified and subsequently dropped from the dataset.
-@Leuz_2003 specify that for taxes payable (`delta_TP`) and short-term debt (`delta_STD`), missing data should be treated as zero, but for other variables, dropping these rows is consistent with financial practices to ensure accuracy.
-
 2.	Derive Operating Cash Flow (CFO) that is indirectly derived from operating income minus accruals:
-
 $$
@@
 $$
-\text{EM1} = \text{Median}\left(\frac{\sigma(\text{Operating Income}_{it})}{\sigma(\text{Operating Cash Flow}_{it})}\right) \quad \text{scaled by lagged total assets}
+\text{EM1} = \text{Median}\left(\frac{\frac{\sigma(\text{Operating Income}_{it})}{\text{Total Assets}_{i(t-1)}}}{\frac{\sigma(\text{Operating Cash Flow}_{it})}{\text{Total Assets}_{i(t-1)}}}\right)
 $$
@@
 ## delete. cell 3 on Descriptive Statistics Table
-```{python}
-#| label: fig-boxplot_smp
-#| fig-cap: "Distribution of Discretionary Accruals over Time"
-#| echo: false
-#| fig-format: svg
-
-results['fig_boxplot_smp']
-```
-
-```{python}
-#| label: fig-scatter_dd_lnta
-#| fig-cap: "Dechow and Dichev DA and Firm Size"
-#| echo: false
-#| fig-format: svg
-
-results['fig_scatter_dd_lnta']
-```
 
@@
 
-```{python}
-#| echo: false
-#| output: asis
-
-lat_tab = rename_var_to_label(results['Correlations'], False).split('\n')
-
-lat_tab[2] = '\\\\[-0.9ex]\\hline \\hline \\\\[-0.9ex]'
-lat_tab[4] = '\\hline\\\\[-1.8ex]'
-lat_tab[12] = '\\\\[-0.9ex]\\hline \\hline \\\\[-0.9ex]'
-lat_tab[14] = '\\begin{tablenotes}[flushleft]'
-
-print('\n'.join([
-    "\\begin{table}[!htbp]",
-    "\centering",
-    "\caption{\label{tab:CorrTable}Correlations}",
-    '\label{tab:corr}',
-    '\n'.join(lat_tab),
-    "\end{table}"
-]))
-```
-
-```{python}
-#| echo: false
-#| output: asis
-
-print('\n'.join([
-    '\\begin{table}[!htbp] \\centering',
-    '\\caption{\\label{tab:reg}Regressions}', 
-    '\label{tab:reg}',
-    rename_var_to_label('\n'.join(results['Regression'].split('\n')[3:-1])),
-    "\end{table}"
-]))
-```
-
 # Results {#sec-results}
@@
 
-The replicated table shows some similarities with the original table by @Leuz_2003 in that . These findings are in line with valuation theory that .
+The replicated table shows some similarities with the original table by @Leuz_2003 in that . These findings are in line with valuation ***theory*** that .
 
@@
 
+The comparison between the calculated EM1 values and those reported in the original paper reveals a strong alignment, with only minor differences observed across most countries.
+The overall pattern of EM1 values is consistent, effectively capturing the underlying trends in earnings management across different countries.
+Any differences, such as those for the Philippines, Taiwan, and Norway, are relatively small and likely due to variations in data sources, updates, or slight methodological differences as outlined in @sec-research_design_assumptions.
+The summary statistics (mean, median, min, max and standard deviation) also closely match those in the original study, further supporting the robustness of the replication process.
+The results indicate that Anglo-American countries tend to have higher EM1 values, suggesting less earnings smoothing compared to Continental Europe and Asia, consistent with the conclusion of @Leuz_2003 that earnings are smoother in Continental Europe and Asia than in Anglo-American countries, after controlling for the volatility of cash flows.
+
 # Conclusion {#sec-conclusion}
 
-##gpt: Summarize the key findings and their significance. take results and insert them here
+##gpt: Summarize the key findings and their significance.
 

### 2024-08-09 code/python/do_analysis.py

@@ -98,6 +98,9 @@ def calculate_em1(df):
     print("EM1 (first 15 rows):")
     print(df[['item6105', 'year_', 'EM1']].head(15))
     
+    # Drop rows where EM1 is NaN
+    df = df.dropna(subset=['EM1'])
+    
     # Group by country (item6026) and take the median of EM1 for each country
     country_em1 = df.groupby('item6026')['EM1'].median().reset_index()
     # Round the EM1 results to three decimal places

### 2024-08-09 doc/paper.qmd
documentation update
added table of contents (the paper is getting longer)

@@
     number-sections: true
-    toc: false
+    toc: true
 fig_caption: yes
@@
 This excludes data reported on a current, daily, or quarterly basis.
-
 After retaining only relevant company data, filtering out financial institutions, and focusing on the specified countries, the processed data was saved to a CSV file in the `data/pulled` directory, as specified in the configuration file `config/pull_data_cfg`​.

### 2024-08-10 doc/paper.qmd
do analysis py

@@
 
-### Smoothing Reported Operating Earnings Using Accruals (EM1)
+### Smoothing Reported Operating Earnings Using Accruals (EM1) {#sec-EM1}
 `EM1` is a measure that assesses how much insiders (e.g., company executives) smooth or stabilize reported earnings by manipulating accruals.
@@
 ### Smoothing and the Correlation Between Changes in Accounting Accruals and Operating Cash Flows (EM2)
+`EM2` is designed to capture the extent to which insiders use accounting discretion to smooth reported earnings by examining the correlation between changes in accruals and changes in operating cash flows.
+A stronger negative correlation between these variables indicates that accruals are being used to buffer the impact of cash flow fluctuations, potentially masking the true economic performance of the firm.
+
+To calculate this measure:
+
+1. Compute Accruals and CFO using same Formulas used in @sec-EM1.
+
+2. Calculate the change in Accruals and CFO for each firm-year:
+$$
+\Delta \text{Accruals}_{it} = \text{Accruals}_{it} - \text{Accruals}_{i(t-1)}
+$$
+$$
+\Delta \text{CFO}_{it} = \text{CFO}_{it} - \text{CFO}_{i(t-1)}
+$$
+
+3. Scale both changes by lagged total assets:
+$$
+\frac{\Delta \text{Accruals}_{it}}{\text{Total Assets}_{i(t-1)}}, \quad \frac{\Delta \text{CFO}_{it}}{\text{Total Assets}_{i(t-1)}}
+$$
+
+4. Calculate EM1, which is computed as the Spearman correlation between these scaled changes, aggregated at the country level:
+$$
+\text{EM2} = \rho\left(\frac{\Delta \text{Accruals}_{it}}{\text{Total Assets}_{i(t-1)}}, \frac{\Delta \text{CFO}_{it}}{\text{Total Assets}_{i(t-1)}}\right)
+$$
 

### 2024-08-10 code/python/do_analysis.py

@@ -3,20 +3,14 @@ import pickle
 import pandas as pd
 
 def main():
-    # Load prepared data
+    # Load the prepared data
     financial_data = load_data()
 
-    # Calculate accruals
-    financial_data = calculate_accruals(financial_data)
-
-    # Calculate Operating Cash Flow (CFO)
-    financial_data = calculate_cfo(financial_data)
-
     # Calculate EM1
-    em1_results = calculate_em1(financial_data)
+    country_em1, summary_stats = calculate_em1(financial_data)
 
     # Save and print results
-    save_results(em1_results)
+    save_results(country_em1, summary_stats)
 
 def load_data():
     """
@@ -25,108 +19,68 @@ def load_data():
     df = pd.read_csv('data/generated/financial_data_prepared.csv')
     return df
 
-def calculate_accruals(df):
+def calculate_em1(df):
     """
-    Calculate accruals using the provided formula.
-    If a firm does not report information on taxes payable or short-term debt,
-    then the change in both variables is assumed to be zero.
+    Calculate EM1 for each firm and then take the country-level median.
+    This function combines all necessary steps into one.
     """
-
+    # Step 1: Calculate Accruals
     # Group by unique firm (item6105) and calculate diffs within each firm
+    # If a firm does not report information on taxes payable or short-term debt,
+    # then the change in both variables is assumed to be zero (per paper).
     df['delta_CA'] = df.groupby('item6105')['item2201'].diff()  # Change in total current assets
     df['delta_Cash'] = df.groupby('item6105')['item2003'].diff()  # Change in cash and cash equivalents
     df['delta_CL'] = df.groupby('item6105')['item3101'].diff()  # Change in total current liabilities
     df['delta_STD'] = df.groupby('item6105')['item3051'].diff().fillna(0)  # Apply fillna only for STD
     df['delta_TP'] = df.groupby('item6105')['item3063'].diff().fillna(0)  # Apply fillna only for TP
-    df['Dep'] = df['item1151']  # No fillna applied to Depreciation and amortization expense
+    df['Dep'] = df['item1151']  # Depreciation and amortization expense
 
     df['Accruals'] = (df['delta_CA'] - df['delta_Cash']) - (df['delta_CL'] - df['delta_STD'] - df['delta_TP']) - df['Dep']
 
-    # Identify rows with NaN in Accruals
-    nan_accruals = df[df['Accruals'].isna()]
-
[…]

### 2024-08-10 code/python/do_analysis.py
EM2

@@ -7,10 +7,13 @@ def main():
     financial_data = load_data()
 
     # Calculate EM1
-    country_em1, summary_stats = calculate_em1(financial_data)
+    country_em1, summary_stats_em1 = calculate_em1(financial_data)
+
+    # Calculate EM2
+    country_em2, summary_stats_em2 = calculate_em2(financial_data)
 
     # Save and print results
-    save_results(country_em1, summary_stats)
+    save_results(country_em1, summary_stats_em1, country_em2, summary_stats_em2)
 
 def load_data():
     """
@@ -65,22 +68,54 @@ def calculate_em1(df):
     country_em1['EM1'] = country_em1['EM1'].round(3)
 
     # Step 8: Calculate summary statistics (mean, median, std, min, max) for EM1 across countries
-    summary_stats = country_em1['EM1'].agg(['mean', 'median', 'std', 'min', 'max']).round(3)
+    summary_stats_em1 = country_em1['EM1'].agg(['mean', 'median', 'std', 'min', 'max']).round(3)
 
-    return country_em1, summary_stats
+    return country_em1, summary_stats_em1
 
-def save_results(country_em1, summary_stats):
+def calculate_em2(df):
     """
-    Save the EM1 results to a pickle file and print them.
+    Calculate EM2, which is the Spearman correlation between the change in Accruals and CFO,
+    both scaled by lagged total assets.
     """
-    with open('output/em1_results.pickle', 'wb') as f:
-        pickle.dump({'country_em1': country_em1, 'summary_stats': summary_stats}, f)
-    
+    # Step 1: Calculate the change in Accruals and CFO
+    df['delta_Accruals'] = df.groupby('item6105')['Accruals'].diff()
+    df['delta_CFO'] = df.groupby('item6105')['CFO'].diff()
+
+    # Step 2: Scale by lagged total assets
+    df['scaled_delta_Accruals'] = df['delta_Accruals'] / df['lagged_total_assets']
+    df['scaled_delta_CFO'] = df['delta_CFO'] / df['lagged_total_assets']
+
+    # Step 3: Calculate EM2 as the Spearman correlation between scaled changes
+    country_em2 = df.groupby('item6026').apply(lambda x: x[['scaled_delta_Accruals', 'scaled_delta_CFO']].corr(method='spearman').iloc[0, 1]).reset_index()
+    country_em2.columns = ['item6026', 'EM2']
+    country_em2['EM2'] = country_em2['EM2'].round(3)
+
[…]

### 2024-08-10 code/python/do_analysis.py
EM2

@@ -81,17 +81,30 @@ def calculate_em2(df):
     df['delta_Accruals'] = df.groupby('item6105')['Accruals'].diff()
     df['delta_CFO'] = df.groupby('item6105')['CFO'].diff()
 
+    # Print the first few rows to check the calculated changes in Accruals and CFO
+    print("Delta Accruals and Delta CFO (first 10 rows):")
+    print(df[['item6105', 'year_', 'Accruals', 'CFO', 'delta_Accruals', 'delta_CFO']].head(10))
+
     # Step 2: Scale by lagged total assets
     df['scaled_delta_Accruals'] = df['delta_Accruals'] / df['lagged_total_assets']
     df['scaled_delta_CFO'] = df['delta_CFO'] / df['lagged_total_assets']
 
+    # Print the first few rows to check the scaled changes
+    print("Scaled Delta Accruals and Scaled Delta CFO (first 10 rows):")
+    print(df[['item6105', 'year_', 'lagged_total_assets', 'scaled_delta_Accruals', 'scaled_delta_CFO']].head(10).to_string(index=False))
+
     # Step 3: Calculate EM2 as the Spearman correlation between scaled changes
     country_em2 = df.groupby('item6026').apply(lambda x: x[['scaled_delta_Accruals', 'scaled_delta_CFO']].corr(method='spearman').iloc[0, 1]).reset_index()
     country_em2.columns = ['item6026', 'EM2']
+    # Round the EM2 results to three decimal places
     country_em2['EM2'] = country_em2['EM2'].round(3)
 
     # Step 4: Calculate summary statistics (mean, median, std, min, max) for EM2 across countries
     summary_stats_em2 = country_em2['EM2'].agg(['mean', 'median', 'std', 'min', 'max']).round(3)
+
+    # Print the summary statistics
+    print("EM2 Summary Statistics:")
+    print(summary_stats_em2)
 
     return country_em2, summary_stats_em2
 

### 2024-08-10 code/python/testEM1.py

@@ -1,0 +1,132 @@
+# We start by loading the libraries that we will use in this analysis.
+import pickle
+import pandas as pd
+
+def main():
+    # Load prepared data
+    financial_data = load_data()
+
+    # Calculate accruals
+    financial_data = calculate_accruals(financial_data)
+
+    # Calculate Operating Cash Flow (CFO)
+    financial_data = calculate_cfo(financial_data)
+
+    # Calculate EM1
+    em1_results = calculate_em1(financial_data)
+
+    # Save and print results
+    save_results(em1_results)
+
+def load_data():
+    """
+    Load the prepared financial data from the previous step.
+    """
+    df = pd.read_csv('data/generated/financial_data_prepared.csv')
+    return df
+
+def calculate_accruals(df):
+    """
+    Calculate accruals using the provided formula.
+    If a firm does not report information on taxes payable or short-term debt,
+    then the change in both variables is assumed to be zero.
+    """
+
+    # Group by unique firm (item6105) and calculate diffs within each firm
+    df['delta_CA'] = df.groupby('item6105')['item2201'].diff()  # Change in total current assets
+    df['delta_Cash'] = df.groupby('item6105')['item2003'].diff()  # Change in cash and cash equivalents
+    df['delta_CL'] = df.groupby('item6105')['item3101'].diff()  # Change in total current liabilities
+    df['delta_STD'] = df.groupby('item6105')['item3051'].diff().fillna(0)  # Apply fillna only for STD
+    df['delta_TP'] = df.groupby('item6105')['item3063'].diff().fillna(0)  # Apply fillna only for TP
+    df['Dep'] = df['item1151']  # No fillna applied to Depreciation and amortization expense
+
+    df['Accruals'] = (df['delta_CA'] - df['delta_Cash']) - (df['delta_CL'] - df['delta_STD'] - df['delta_TP']) - df['Dep']
+
+    # Identify rows with NaN in Accruals
+    nan_accruals = df[df['Accruals'].isna()]
+
+    # Print the first 10 rows with NaN in Accruals
+    print("First 10 rows with NaN in Accruals:")
+    print(nan_accruals.head(10))
+
+    # Count and print the number of NaN rows in Accruals
+    nan_count = df['Accruals'].isna().sum()
+    print(f"Number of rows with NaN in Accruals: {nan_count}")
+
+    # Print the first few rows to check the calculated accruals
+    print(df[['item6105', 'year_', 'Accruals']].head(10))
[…]

### 2024-08-10 code/python/testEM2.py

@@ -1,0 +1,134 @@
+# We start by loading the libraries that we will use in this analysis.
+import pickle
+import pandas as pd
+
+def main():
+    # Load the prepared data
+    financial_data = load_data()
+
+    # Calculate EM1
+    country_em1, summary_stats_em1 = calculate_em1(financial_data)
+
+    # Calculate EM2
+    country_em2, summary_stats_em2 = calculate_em2(financial_data)
+
+    # Save and print results
+    save_results(country_em1, summary_stats_em1, country_em2, summary_stats_em2)
+
+def load_data():
+    """
+    Load the prepared financial data from the previous step.
+    """
+    df = pd.read_csv('data/generated/financial_data_prepared.csv')
+    return df
+
+def calculate_em1(df):
+    """
+    Calculate EM1 for each firm and then take the country-level median.
+    This function combines all necessary steps into one.
+    """
+    # Step 1: Calculate Accruals
+    # Group by unique firm (item6105) and calculate diffs within each firm
+    # If a firm does not report information on taxes payable or short-term debt,
+    # then the change in both variables is assumed to be zero (per paper).
+    df['delta_CA'] = df.groupby('item6105')['item2201'].diff()  # Change in total current assets
+    df['delta_Cash'] = df.groupby('item6105')['item2003'].diff()  # Change in cash and cash equivalents
+    df['delta_CL'] = df.groupby('item6105')['item3101'].diff()  # Change in total current liabilities
+    df['delta_STD'] = df.groupby('item6105')['item3051'].diff().fillna(0)  # Apply fillna only for STD
+    df['delta_TP'] = df.groupby('item6105')['item3063'].diff().fillna(0)  # Apply fillna only for TP
+    df['Dep'] = df['item1151']  # Depreciation and amortization expense
+
+    df['Accruals'] = (df['delta_CA'] - df['delta_Cash']) - (df['delta_CL'] - df['delta_STD'] - df['delta_TP']) - df['Dep']
+
+    # Step 2: Calculate Operating Cash Flow (CFO) by subtracting accruals from operating income.
+    # Handle NaN values where necessary.
+    df['CFO'] = df['item1250'] - df['Accruals']
+
+    # Step 3: Calculate the standard deviations for Operating Income and CFO for each firm (grouped by item1605)
[…]

### 2024-08-10 doc/paper.qmd
EM2
do analysis

@@
 
-4. Calculate EM1, which is computed as the Spearman correlation between these scaled changes, aggregated at the country level:
+4. Calculate EM2, which is computed as the Spearman correlation between these scaled changes, aggregated at the country level:
 $$
@@
 
+The replication findings regarding EM2 indicate that earnings smoothing is more pronounced in countries like Greece and Japan, while countries like Canada and the United States exhibit less pervasive smoothing behavior, in line with the findings of @Leuz_2003.
+Despite some differences in magnitude, the overall trends are consistent, reinforcing the idea that earnings smoothing varies significantly across different countries.
+In fact, these small variations do not significantly alter the overall conclusions.
+Specifically, the minimum and maximum EM2 values in the replication closely match the original findings, with Greece showing the most earnings smoothing (min: -0.942 vs.
+-0.928) and Norway the least (max: -0.786 vs.
+-0.722).
+This consistency reinforces the accuracy of the replication.
+
 # Conclusion {#sec-conclusion}

### 2024-08-10 code/python/do_analysis.py

@@ -41,10 +41,9 @@ def calculate_em1(df):
     df['Accruals'] = (df['delta_CA'] - df['delta_Cash']) - (df['delta_CL'] - df['delta_STD'] - df['delta_TP']) - df['Dep']
 
     # Step 2: Calculate Operating Cash Flow (CFO) by subtracting accruals from operating income.
-    # Handle NaN values where necessary.
     df['CFO'] = df['item1250'] - df['Accruals']
 
-    # Step 3: Calculate the standard deviations for Operating Income and CFO for each firm (grouped by item1605)
+    # Step 3: Calculate the standard deviations for Operating Income and CFO for each firm
     df['std_operating_income'] = df.groupby('item6105')['item1250'].transform('std')
     df['std_cfo'] = df.groupby('item6105')['CFO'].transform('std')
 
@@ -81,30 +80,20 @@ def calculate_em2(df):
     df['delta_Accruals'] = df.groupby('item6105')['Accruals'].diff()
     df['delta_CFO'] = df.groupby('item6105')['CFO'].diff()
 
-    # Print the first few rows to check the calculated changes in Accruals and CFO
-    print("Delta Accruals and Delta CFO (first 10 rows):")
-    print(df[['item6105', 'year_', 'Accruals', 'CFO', 'delta_Accruals', 'delta_CFO']].head(10))
-
     # Step 2: Scale by lagged total assets
     df['scaled_delta_Accruals'] = df['delta_Accruals'] / df['lagged_total_assets']
     df['scaled_delta_CFO'] = df['delta_CFO'] / df['lagged_total_assets']
 
-    # Print the first few rows to check the scaled changes
-    print("Scaled Delta Accruals and Scaled Delta CFO (first 10 rows):")
-    print(df[['item6105', 'year_', 'lagged_total_assets', 'scaled_delta_Accruals', 'scaled_delta_CFO']].head(10).to_string(index=False))
-
     # Step 3: Calculate EM2 as the Spearman correlation between scaled changes
-    country_em2 = df.groupby('item6026').apply(lambda x: x[['scaled_delta_Accruals', 'scaled_delta_CFO']].corr(method='spearman').iloc[0, 1]).reset_index()
+    country_em2 = df.groupby('item6026').apply(
+        lambda x: x[['scaled_delta_Accruals', 'scaled_delta_CFO']].corr(method='spearman').iloc[0, 1]
+    ).reset_index()
     country_em2.columns = ['item6026', 'EM2']
[…]

### 2024-08-10 doc/paper.qmd
em3

@@
 ### Smoothing Reported Operating Earnings Using Accruals (EM1) {#sec-EM1}
-`EM1` is a measure that assesses how much insiders (e.g., company executives) smooth or stabilize reported earnings by manipulating accruals.
+EM1 is a measure that assesses how much insiders (e.g., company executives) smooth or stabilize reported earnings by manipulating accruals.
 The concept is that firms may attempt to reduce fluctuations in their reported operating income to present a more stable financial performance to investors.
@@
 ### Smoothing and the Correlation Between Changes in Accounting Accruals and Operating Cash Flows (EM2)
-`EM2` is designed to capture the extent to which insiders use accounting discretion to smooth reported earnings by examining the correlation between changes in accruals and changes in operating cash flows.
+EM2 is designed to capture the extent to which insiders use accounting discretion to smooth reported earnings by examining the correlation between changes in accruals and changes in operating cash flows.
 A stronger negative correlation between these variables indicates that accruals are being used to buffer the impact of cash flow fluctuations, potentially masking the true economic performance of the firm.
@@
 
-4. Calculate EM2, which is computed as the Spearman correlation between these scaled changes, aggregated at the country level:
+4.
+Calculate EM2, which is computed as the Spearman correlation between these scaled changes, aggregated at the country level over the pooled set of firms within each country, meaning that the correlation reflects the aggregated relationship between changes in accruals and changes in CFO across all firms in that country:
 $$
@@
 ### Discretion in Reported Earnings: The Magnitude of Accruals (EM3)
+EM3 is a measure that obtains the degree of discretion that insiders exercise when reporting earnings, particularly when it comes to misstating a firm’s economic performance.
+It focuses on the magnitude of accruals as a proxy for earnings management.
+Insiders can use accruals to overstate or understate earnings, thereby achieving certain financial targets or presenting a skewed view of the firm’s performance.
+
+To calculate this measure:
+
+1. Compute Accruals and CFO using same Formulas used in @sec-EM1.
+
+2. Calculate the absolute values of accruals and CFO for each firm-year.
+
+3.
+EM3 is calculated as the country-level median ratio of the absolute value of firms’ accruals to the absolute value of their cash flow from operations (CFO).
+This ratio is used as it scales the magnitude of accruals to a measure of the firm’s size and performance, allowing for comparison across firms of different sizes and performance levels:
+$$
+\text{EM3} = \text{Median}\left(\frac{|\text{Accruals}_{it}|}{|\text{CFO}_{it}|}\right)
+$$
 

### 2024-08-10 code/python/do_analysis.py

@@ -12,8 +12,11 @@ def main():
     # Calculate EM2
     country_em2, summary_stats_em2 = calculate_em2(financial_data)
 
+    # Calculate EM3
+    country_em3, summary_stats_em3 = calculate_em3(financial_data)
+
     # Save and print results
-    save_results(country_em1, summary_stats_em1, country_em2, summary_stats_em2)
+    save_results(country_em1, summary_stats_em1, country_em2, summary_stats_em2, country_em3, summary_stats_em3)
 
 def load_data():
     """
@@ -48,7 +51,6 @@ def calculate_em1(df):
     df['std_cfo'] = df.groupby('item6105')['CFO'].transform('std')
 
     # Step 4: Retrieve Lagged Total Assets
-    # For scaling, identify the total assets from the prior year (making use of year 0 as well. Otherwise would have dropped it in accruals step)
     df['lagged_total_assets'] = df.groupby('item6105')['item2999'].shift(1)
 
     # Step 5: Scale the standard deviations individually by lagged total assets
@@ -57,9 +59,6 @@ def calculate_em1(df):
 
     # Step 6: Calculate EM1 as the ratio of the scaled standard deviations
     df['EM1'] = df['scaled_std_operating_income'] / df['scaled_std_cfo']
-
-    # Drop rows where EM1 is NaN
-    ## df = df.dropna(subset=['EM1'])
 
     # Step 7: Group by country (item6026) and calculate the median of EM1 for each country
     country_em1 = df.groupby('item6026')['EM1'].median().reset_index()
@@ -97,16 +96,51 @@ def calculate_em2(df):
 
     return country_em2, summary_stats_em2
 
-def save_results(country_em1, summary_stats_em1, country_em2, summary_stats_em2):
+def calculate_em3(df):
     """
-    Save the EM1 and EM2 results to a pickle file and print them.
+    Calculate EM3, which is the country’s median ratio of the absolute value of accruals
+    and the absolute value of the cash flow from operations.
+    """
+    # Step 1: Calculate the absolute values of Accruals and CFO
+    df['abs_Accruals'] = df['Accruals'].abs()
+    df['abs_CFO'] = df['CFO'].abs()
+
+    # Print the first few rows to check the calculated absolute values
+    print("Absolute Values of Accruals and CFO (first 10 rows):")
+    print(df[['item6105', 'year_', 'Accruals', 'CFO', 'abs_Accruals', 'abs_CFO']].head(10))
+
[…]

### 2024-08-10 doc/paper.qmd
em3

@@
 Insiders can use accruals to overstate or understate earnings, thereby achieving certain financial targets or presenting a skewed view of the firm’s performance.
+Higher scores of EM3 indicate more earnings management, as larger magnitudes of accruals relative to cash flow from operations suggest greater discretion in financial reporting.
 
@@
 
-The replicated table shows some similarities with the original table by @Leuz_2003 in that . These findings are in line with valuation ***theory*** that .
+##The replicated table shows some similarities with the original table by @Leuz_2003 in that . These findings are in line with valuation ***theory*** that .
 
-However, the patterns are less clear in the replicated table (in particular ), which stands in contrast to @Leuz_2003.
+##However, the patterns are less clear in the replicated table (in particular ), which stands in contrast to @Leuz_2003.
 
@@
 This consistency reinforces the accuracy of the replication.
+
+The replicated EM3 results are very similar to those of the original study, with close alignment in summary statistics and consistent identification of South Africa and Germany as the countries with the minimum and maximum EM3 values.
+Additionally, the replication confirms that the magnitude of firms’ accruals relative to their operating cash flows is smaller in the U.K.
+and the U.S.
+compared to Austria, Germany, and South Korea, consistent with the original findings.
 

### 2024-08-10 code/python/do_analysis.py

@@ -105,24 +105,14 @@ def calculate_em3(df):
     df['abs_Accruals'] = df['Accruals'].abs()
     df['abs_CFO'] = df['CFO'].abs()
 
-    # Print the first few rows to check the calculated absolute values
-    print("Absolute Values of Accruals and CFO (first 10 rows):")
-    print(df[['item6105', 'year_', 'Accruals', 'CFO', 'abs_Accruals', 'abs_CFO']].head(10))
-
     # Step 2: Calculate the ratio of abs(Accruals) to abs(CFO)
-    df['Ratio'] = df['abs_Accruals'] / df['abs_CFO']
+    df['EM3'] = df['abs_Accruals'] / df['abs_CFO']
 
-    # Print the first few rows to check the calculated Ratio
-    print("Ratio of Absolute Accruals to Absolute CFO (first 10 rows):")
-    print(df[['item6105', 'year_', 'abs_Accruals', 'abs_CFO', 'Ratio']].head(10))
+    # Step 3: Group by country and calculate the median of EM3 for each country
+    country_em3 = df.groupby('item6026')['EM3'].median().reset_index()
 
-    # Step 3: Group by country and calculate the median of the Ratio for each country
-    country_em3 = df.groupby('item6026')['Ratio'].median().reset_index()
-    country_em3['EM3'] = country_em3['Ratio'].round(3)
-
-    # Print the country-level EM3 results
-    print("Country-Level EM3 Results (first 10 rows):")
-    print(country_em3.head(10))
+    # Round the EM3 results to three decimal places
+    country_em3['EM3'] = country_em3['EM3'].round(3)
 
     # Step 4: Calculate summary statistics (mean, median, std, min, max) for EM3 across countries
     summary_stats_em3 = country_em3['EM3'].agg(['mean', 'median', 'std', 'min', 'max']).round(3)

### 2024-08-10 doc/paper.qmd
em3

@@
 
-## Step 2: Data Preparation
+## Step 2: Data Preparation {#sec-step2_preparation}
 To verify the pulled data, I checked the dataset for duplicate firm-year observations based on the combination of the Worldscope Permanent ID (`item6105`) and the fiscal year (`year_`).
@@
 ### Discretion in Reported Earnings: Small Loss Avoidance (EM4)
+EM4 captures the extent to which firms manage their reported earnings to avoid small losses.
+This is an important measure that quantifies the propensity of firms to avoid small losses through earnings management.
+By calculating the ratio of small profits to small losses, it provides a clear indication of how widespread this practice is within a particular country.
+A higher EM4 value suggests that firms in a given country are more likely to use accounting discretion to convert small losses into small profits, reflecting a tendency toward earnings management.
+Note that I used `item1651` (Net Income before Preferred Dividends) as the variable for net earnings, as mentioned in @sec-step2_preparation.
+
+EM4 is calculated as the ratio of the number of small profits to the number of small losses for each country, where small profits and small losses are determined by scaling net earnings by lagged total assets:
+$$
+\begin{aligned}
+\text{EM4}&= \frac{\text{Small Profits}}{\text{Small Losses}} \\
+& \\
+\text{where:} \\
+\text{Small Profits} & : \frac{\text{Net Earnings}_{it}}{\text{Total Assets}_{it-1}} \in [0, 0.01] \\
+\text{Small Losses}  & : \frac{\text{Net Earnings}_{it}}{\text{Total Assets}_{it-1}} \in [-0.01, 0) \\
+\end{aligned}
+$$
 

### 2024-08-10 doc/paper.qmd
em4

@@
 ### Aggregate Measure of Earnings Management
+The aggregate earnings management score is a summary measure that captures the overall level of earnings management in each country.
+This measure is calculated by ranking countries on each of the four individual earnings management measures (EM1-EM4), where a higher rank indicates a greater degree of earnings management.
+The final aggregate score is the average of these ranks, providing a comprehensive view of a country’s earnings management practices.
+
+To calculate this measure:
+$$
+\text{Aggregate Earnings Management Score}_{\text{country}} = \frac{\text{Rank}_{\text{EM1}} + \text{Rank}_{\text{EM2}} + \text{Rank}_{\text{EM3}} + \text{Rank}_{\text{EM4}}}{4}
+$$
 
@@
 
+The replicated results show significant regional differences in EM4 values, reflecting varying degrees of loss avoidance.
+Spain has the highest EM4 value, highlighting strong loss avoidance among European firms.
+Norway, with the lowest EM4 value, corresponds to the minimum value in the original paper, confirming less aggressive loss avoidance.
+High EM4 values in Asian countries like Indonesia (5.667) and India (6.000) further support the trend that European and Asian firms exhibit greater loss avoidance than their Anglo-American counterparts, such as the United States (1.912) and Canada (2.402), which have lower EM4 values.
+These results align with the original study's conclusions on regional loss avoidance behavior.
+
 # Conclusion {#sec-conclusion}

### 2024-08-10 code/python/do_analysis.py

@@ -14,9 +14,12 @@ def main():
 
     # Calculate EM3
     country_em3, summary_stats_em3 = calculate_em3(financial_data)
+    
+    # Calculate EM4
+    country_em4, summary_stats_em4 = calculate_em4(financial_data)
 
     # Save and print results
-    save_results(country_em1, summary_stats_em1, country_em2, summary_stats_em2, country_em3, summary_stats_em3)
+    save_results(country_em1, summary_stats_em1, country_em2, summary_stats_em2, country_em3, summary_stats_em3, country_em4, summary_stats_em4)
 
 def load_data():
     """
@@ -119,9 +122,34 @@ def calculate_em3(df):
 
     return country_em3, summary_stats_em3
 
-def save_results(country_em1, summary_stats_em1, country_em2, summary_stats_em2, country_em3, summary_stats_em3):
+def calculate_em4(df):
     """
-    Save the EM1, EM2, and EM3 results to a pickle file and print them.
+    Calculate EM4, which is the ratio of the number of small profits to the number of small losses for each country.
+    Small profits and losses are defined based on net earnings (item1651) scaled by lagged total assets.
+    """
+    # Step 1: Calculate Net Earnings scaled by Lagged Total Assets
+    df['scaled_net_earnings'] = df['item1651'] / df['lagged_total_assets']
+
+    # Step 2: Identify Small Profits and Small Losses
+    df['small_profits'] = ((df['scaled_net_earnings'] >= 0) & (df['scaled_net_earnings'] <= 0.01)).astype(int)
+    df['small_losses'] = ((df['scaled_net_earnings'] >= -0.01) & (df['scaled_net_earnings'] < 0)).astype(int)
+
+    # Step 3: Calculate EM4 as the ratio of Small Profits to Small Losses for each country
+    country_em4 = df.groupby('item6026').apply(
+        lambda x: x['small_profits'].sum() / max(1, x['small_losses'].sum())  # Avoid division by zero
+    ).reset_index()
+    country_em4.columns = ['item6026', 'EM4']
+    # Round the EM4 results to three decimal places
+    country_em4['EM4'] = country_em4['EM4'].round(3)
+
+    # Step 4: Calculate summary statistics (mean, median, std, min, max) for EM4 across countries
+    summary_stats_em4 = country_em4['EM4'].agg(['mean', 'median', 'std', 'min', 'max']).round(3)
+
+    return country_em4, summary_stats_em4
+
[…]

### 2024-08-10 doc/paper.qmd
update results section

@@
 
+Table 1 provides descriptive statistics for the four individual earnings management measures as well as the aggregate earnings management score, where countries are sorted in descending order based on their aggregate score.
+The discussion on replication findings is presented in @sec-results.
+
 ## delete from fikir.
@@
 Given that the number of observations in this project is significantly higher than in the original study, Table 1 from @Leuz_2003 was partially replicated (focusing only on columns related to countries and firm-year observations) to highlight specific discrepancies that could arise for certain countries.
-The results in the table 2 below show a higher number of firm-year observations for almost all countries compared to the paper by @Leuz_2003, likely due to database updates leading to a more comprehensive dataset.
+The results in the Table 2 below show a higher number of firm-year observations for almost all countries compared to the paper by @Leuz_2003, likely due to database updates leading to a more comprehensive dataset.
 The trend in the number of firm-year observations across countries is consistent with the original study, with larger capital markets like Japan, the United Kingdom, and the United States having the highest numbers.
@@
 The final aggregate score is the average of these ranks, providing a comprehensive view of a country’s earnings management practices.
+The countries are sorted in descending order based on their aggregate score.
 
@@
 
-## delete. cell 3 on Descriptive Statistics Table
-
-```{python}
-#| echo: false
-#| output: asis
-
-lat_tab = rename_var_to_label(results['Descriptive Statistics']).split('\n')
-lat_tab[4] = '\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]'
-lat_tab[6] = '\\hline\\\\[-1.8ex]'
-lat_tab[-4] = '\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]'
-
-print('\n'.join([
-    '\\begin{table}[!htbp]',  
-    '\caption{\label{tab:descriptives}Descriptive Statistics}',
-    '\label{tab:tab-descriptives-sample}',
-    '\\begin{adjustbox}{width=\\textwidth}',
-    '\\begin{threeparttable}',
-    '\n'.join(lat_tab[3:-2]),
-    '\\begin{tablenotes}[flushleft]',
-    '\setlength{\labelsep}{0pt}',
-    '\\footnotetext',
-    f'\\item Note: The data is obtained from the Compustat U.S. as provided by WRDS. The sample covers the period {int(min_fyear)} to {int(max_fyear)} and {no_unique_firms:,} unique firms.',
-    '\end{tablenotes}',
-    '\end{threeparttable}',
-    '\end{adjustbox}',
-    '\\end{table}',
-]))
-```
-
 # Results {#sec-results}
-
-##Cross-check your results with the original table to ensure accuracy.
-Compare your results with the results in the paper.
-If you notice differences, discuss where they may originate.
-Make sure to be as specific as possible.
-This could involve looking at the data selection, calculation methods, or any assumptions that might differ from the original; changes in the data over time, different interpretations of the methodology, or limitations in the data sources used.
-Consider whether the differences are significant and what they might mean for the conclusions drawn in the paper.
-
-##The replicated table shows some similarities with the original table by @Leuz_2003 in that . These findings are in line with valuation ***theory*** that .
-
-##However, the patterns are less clear in the replicated table (in particular ), which stands in contrast to @Leuz_2003.
-
 The comparison between the calculated EM1 values and those reported in the original paper reveals a strong alignment, with only minor differences observed across most countries.
@@
 
+The comparison between the original and replicated aggregate earnings management scores reveals a strong overall alignment with consistent key trends.
+In the original study, Austria and Greece are at the top of the rankings with aggregate scores of 28.3, indicating the highest levels of earnings management.
+In the replication, Greece and Austria remain at the top with slightly adjusted scores of 27.5 and 27.2, respectively.
+Taiwan and South Korea also rank high, mirroring their positions in the original study.
+While some differences in specific rankings and scores exist, such as the United States consistently having the lowest level of earnings management (2.0 in the original and 3.5 in the replication), these variations are minor and do not spoil overall consistency.
+
+In summary, the replication effectively captures the original study’s key findings, confirming that European and Asian countries generally exhibit higher levels of earnings management than their Anglo-American counterparts.
+The strong similarity between the original and replicated results underscores the robustness and reliability of the findings.
+
 # Conclusion {#sec-conclusion}
@@
 
+The replicated results show significant differences in earnings management practices across countries, with consistent patterns in the relative magnitudes of the individual measures.
+These differences are particularly evident in areas such as loss avoidance and earnings smoothing.
+In line with the original study, the replication suggests that firms in countries with more developed equity markets, dispersed ownership structures, strong investor rights, and robust legal enforcement tend to engage in less earnings management.
+This consistent finding across both the original and replicated studies reinforces the idea that the institutional environment plays a critical role in shaping corporate earnings management behaviors on an international scale.
+
+The replicated analysis also reveals that the individual earnings management measures are highly correlated, leading to similar rankings across the four measures and the aggregate earnings management score.
+This consistency supports the original study’s finding that a single underlying factor likely represents the four individual measures.
+In line with the original results, the replicated country rankings based on the aggregate score show high ranks for countries like Greece, Austria, and South Korea, and lower ranks for countries such as Australia, the U.K., and the U.S.
+This reinforces the robustness of the aggregated measure in capturing the overall level of earnings management across different countries.
+
 This final assignment has required the comprehensive application of all skills learned and feedback received from previous assignments, making it a more complex project workflow.

### 2024-08-10 code/python/do_analysis.py

@@ -18,8 +18,11 @@ def main():
     # Calculate EM4
     country_em4, summary_stats_em4 = calculate_em4(financial_data)
 
+    # Calculate Aggregate Earnings Management Score
+    aggregate_em = calculate_aggregate(country_em1, country_em2, country_em3, country_em4)
+
     # Save and print results
-    save_results(country_em1, summary_stats_em1, country_em2, summary_stats_em2, country_em3, summary_stats_em3, country_em4, summary_stats_em4)
+    save_results(country_em1, summary_stats_em1, country_em2, summary_stats_em2, country_em3, summary_stats_em3, country_em4, summary_stats_em4, aggregate_em)
 
 def load_data():
     """
@@ -147,9 +150,51 @@ def calculate_em4(df):
 
     return country_em4, summary_stats_em4
 
-def save_results(country_em1, summary_stats_em1, country_em2, summary_stats_em2, country_em3, summary_stats_em3, country_em4, summary_stats_em4):
+def calculate_aggregate(df_em1, df_em2, df_em3, df_em4):
     """
-    Save the EM1, EM2, EM3, and EM4 results to a pickle file and print them.
+    Calculate the aggregate earnings management score for each country.
+    This function averages the ranks of EM1, EM2, EM3, and EM4.
+    """
+    # Step 1: Merge the EM1, EM2, EM3, and EM4 dataframes on the 'item6026' column (country identifier)
+    combined_df = pd.merge(df_em1, df_em2, on='item6026')
+    print("\nAfter merging EM1 and EM2:")
+    print(combined_df.head())
+
+    combined_df = pd.merge(combined_df, df_em3, on='item6026')
+    print("\nAfter merging EM3:")
+    print(combined_df.head())
+
+    combined_df = pd.merge(combined_df, df_em4, on='item6026')
+    print("\nAfter merging EM4:")
+    print(combined_df.head())
+
+    # Step 2: Rank each country by EM1, EM2, EM3, and EM4
+    # Adjusting ranking based on whether higher scores imply more (+) or less (-) earnings management
+    combined_df['Rank_EM1'] = combined_df['EM1'].rank(ascending=False)  # "-" indicates less EM with higher scores
+    combined_df['Rank_EM2'] = combined_df['EM2'].rank(ascending=False)  # "-" indicates less EM with higher scores
[…]

### 2024-08-10 doc/paper.qmd
results section

@@
 # Introduction {#sec-introduction}
-
 The aim of this paper is to illustrate the use of open science tools in empirical accounting research.
@@
 # Research Design Choices and Assumptions {#sec-research_design_assumptions}
-
 The aim of Assignment III is to replicate a specific empirical table (Table 2 Panel A) from the seminal paper by @Leuz_2003.
@@
 The overall pattern of EM1 values is consistent, effectively capturing the underlying trends in earnings management across different countries.
-Any differences, such as those for the Philippines, Taiwan, and Norway, are relatively small and likely due to variations in data sources, updates, or slight methodological differences as outlined in @sec-research_design_assumptions.
-The summary statistics (mean, median, min, max and standard deviation) also closely match those in the original study, further supporting the robustness of the replication process.
-The results indicate that Anglo-American countries tend to have higher EM1 values, suggesting less earnings smoothing compared to Continental Europe and Asia, consistent with the conclusion of @Leuz_2003 that earnings are smoother in Continental Europe and Asia than in Anglo-American countries, after controlling for the volatility of cash flows.
+Any differences, such as those for the Philippines, Taiwan, and Norway, are relatively small and likely due to variations in data sources, updates, or slight methodological differences, as outlined in @sec-research_design_assumptions.
+The summary statistics (mean, median, min, max, and standard deviation) also closely match those in the original study, further supporting the robustness of the replication process.
+The results indicate that Anglo-American countries tend to have higher EM1 values, suggesting less earnings smoothing compared to Continental Europe and Asia, consistent with the conclusion of @Leuz_2003 that earnings are smoother in Continental Europe and Asia than in Anglo-American countries after controlling for the volatility of cash flows.
 
@@
 The replicated EM3 results are very similar to those of the original study, with close alignment in summary statistics and consistent identification of South Africa and Germany as the countries with the minimum and maximum EM3 values.
-Additionally, the replication confirms that the magnitude of firms’ accruals relative to their operating cash flows is smaller in the U.K.
+Additionally, the replication confirms that the magnitude of firms' accruals relative to their operating cash flows is smaller in the U.K.
 and the U.S.
@@
 Taiwan and South Korea also rank high, mirroring their positions in the original study.
-While some differences in specific rankings and scores exist, such as the United States consistently having the lowest level of earnings management (2.0 in the original and 3.5 in the replication), these variations are minor and do not spoil overall consistency.
+While some differences in specific rankings and scores exist, such as the United States consistently having the lowest level of earnings management (2.0 in the original and 3.5 in the replication), these variations are minor and do not undermine the overall consistency.
 
-In summary, the replication effectively captures the original study’s key findings, confirming that European and Asian countries generally exhibit higher levels of earnings management than their Anglo-American counterparts.
+In summary, the replication effectively captures the original study's key findings, confirming that European and Asian countries generally exhibit higher levels of earnings management than their Anglo-American counterparts.
 The strong similarity between the original and replicated results underscores the robustness and reliability of the findings.
@@
 # Conclusion {#sec-conclusion}
-
 ##gpt: Summarize the key findings and their significance.
@@
 
-The analysis revealed that
-
-The replication yields mixed results. While
-
 The replicated results show significant differences in earnings management practices across countries, with consistent patterns in the relative magnitudes of the individual measures.
@@
 
-The replicated analysis also reveals that the individual earnings management measures are highly correlated, leading to similar rankings across the four measures and the aggregate earnings management score.
+The replication analysis also reveals that the individual earnings management measures are highly correlated, leading to similar rankings across the four measures and the aggregate earnings management score.
 This consistency supports the original study’s finding that a single underlying factor likely represents the four individual measures.

### 2024-08-10 doc/paper.qmd
final table

@@
 # Conclusion {#sec-conclusion}
-##gpt: Summarize the key findings and their significance.
+The project effectively demonstrates the use of an open science and collaborative workflow for analyzing earnings management measures across various countries.
+By following a systematic approach and utilizing the TRR 266 Template for Reproducible Empirical Accounting Research, I successfully replicated an empirical table from @Leuz_2003 and provided insights into trends in earnings management practices, which, similar to the original study, vary significantly across countries.
 
-This project demonstrates the effectiveness of using an open science and collaborative workflow for analyzing earnings management measures across various countries.
-By following a step-by-step approach and using the TRR 266 Template for Reproducible Empirical Accounting Research, I was able to replicate an empirical table from @Leuz_2003 and provide insights into the trends of earnings management practices, which vary significantly across countries, similar to @Leuz_2003.
+@Leuz_2003 document systematic differences in the level of earnings management across 31 countries, suggesting that insiders use earnings management (EM) to conceal firm performance and protect their private control benefits.
+The study further argues that stronger investor protection reduces earnings management by limiting insiders’ ability to secure private benefits.
+By conducting a descriptive cluster analysis following the descriptive statistics replicated in this project, the paper identifies groupings of countries with similar institutional characteristics, demonstrating that earnings management varies systematically across these institutional clusters.
 
-The study by @Leuz_2003 documents systematic differences in the level of earnings management across 31 countries.
-It suggests that insiders use EM to conceal firm performance and protect their private control benefits and that stronger investor protection reduces earnings management by limiting insiders’ ability to gain private benefits.
-By conducting a descriptive cluster analysis after descriptive statistics that was replicated in this project, the paper identifies groupings of countries with similar institutional characteristics and demonstrates that earnings management varies systematically across these institutional clusters.
+The replicated results reveal significant differences in earnings management practices across countries, with consistent patterns in the relative magnitudes of the individual measures.
+These differences are particularly notable in areas such as loss avoidance and earnings smoothing.
+In line with the original study, the replication suggests that firms in countries with more developed equity markets, dispersed ownership structures, strong investor rights, and robust legal enforcement tend to engage in less earnings management.
+This consistent finding across both the original and replicated studies underscores the critical role of the institutional environment in shaping corporate earnings management behaviors on an international scale.
 
-The replicated results show significant differences in earnings management practices across countries, with consistent patterns in the relative magnitudes of the individual measures.
-These differences are particularly evident in areas such as loss avoidance and earnings smoothing.
-In line with the original study, the replication suggests that firms in countries with more developed equity markets, dispersed ownership structures, strong investor rights, and robust legal enforcement tend to engage in less earnings management.
-This consistent finding across both the original and replicated studies reinforces the idea that the institutional environment plays a critical role in shaping corporate earnings management behaviors on an international scale.
+The replication analysis also reveals that the individual earnings management measures are highly correlated, resulting in similar rankings across the four measures and the aggregate earnings management score.
+This consistency supports the original study’s conclusion that a single underlying factor likely represents the four individual measures.
+As in the original results, the replicated country rankings based on the aggregate score show high ranks for countries like Greece, Austria, and South Korea, and lower ranks for countries such as Australia, the U.K., and the U.S.
+This reinforces the robustness of the aggregate measure in capturing the overall level of earnings management across different countries.
 
-The replication analysis also reveals that the individual earnings management measures are highly correlated, leading to similar rankings across the four measures and the aggregate earnings management score.
-This consistency supports the original study’s finding that a single underlying factor likely represents the four individual measures.
-In line with the original results, the replicated country rankings based on the aggregate score show high ranks for countries like Greece, Austria, and South Korea, and lower ranks for countries such as Australia, the U.K., and the U.S.
-This reinforces the robustness of the aggregated measure in capturing the overall level of earnings management across different countries.
-
-This final assignment has required the comprehensive application of all skills learned and feedback received from previous assignments, making it a more complex project workflow.
+This final assignment required the comprehensive application of all the skills learned and feedback received from previous assignments, making it a more complex project workflow.
 In the future, this repository can be cloned or forked (if made public) to kickstart further projects on earnings management measures analysis.

### 2024-08-10 code/python/do_analysis.py

@@ -157,16 +157,8 @@ def calculate_aggregate(df_em1, df_em2, df_em3, df_em4):
     """
     # Step 1: Merge the EM1, EM2, EM3, and EM4 dataframes on the 'item6026' column (country identifier)
     combined_df = pd.merge(df_em1, df_em2, on='item6026')
-    print("\nAfter merging EM1 and EM2:")
-    print(combined_df.head())
-
     combined_df = pd.merge(combined_df, df_em3, on='item6026')
-    print("\nAfter merging EM3:")
-    print(combined_df.head())
-
     combined_df = pd.merge(combined_df, df_em4, on='item6026')
-    print("\nAfter merging EM4:")
-    print(combined_df.head())
 
     # Step 2: Rank each country by EM1, EM2, EM3, and EM4
     # Adjusting ranking based on whether higher scores imply more (+) or less (-) earnings management
@@ -175,27 +167,37 @@ def calculate_aggregate(df_em1, df_em2, df_em3, df_em4):
     combined_df['Rank_EM3'] = combined_df['EM3'].rank(ascending=True)  # "+" indicates more EM with higher scores
     combined_df['Rank_EM4'] = combined_df['EM4'].rank(ascending=True)  # "+" indicates more EM with higher scores
 
-    print("\nAfter ranking countries:")
-    print(combined_df[['item6026', 'Rank_EM1', 'Rank_EM2', 'Rank_EM3', 'Rank_EM4']].head(32))
-
     # Step 3: Calculate the aggregate score by averaging the ranks and round to one decimal place
     combined_df['Aggregate_EM_Score'] = combined_df[['Rank_EM1', 'Rank_EM2', 'Rank_EM3', 'Rank_EM4']].mean(axis=1).round(1)
 
-    print("\nAfter calculating the Aggregate EM Score:")
-    print(combined_df[['item6026', 'Aggregate_EM_Score']].head())
+    # Keep all relevant columns for the final table
+    final_table = combined_df[['item6026', 'EM1', 'EM2', 'EM3', 'EM4', 'Aggregate_EM_Score']].sort_values(by='Aggregate_EM_Score', ascending=False).reset_index(drop=True)
+    
+    return final_table
 
-    # Keep only the necessary columns for the result
-    aggregate_results = combined_df[['item6026', 'Aggregate_EM_Score']].sort_values(by='Aggregate_EM_Score', ascending=False).reset_index(drop=True)
+def create_final_table_with_stats(final_table):
+    """
[…]

### 2024-08-10 code/python/do_analysis.py
aggregate

@@ -175,28 +175,36 @@ def calculate_aggregate(df_em1, df_em2, df_em3, df_em4):
     
     return final_table
 
-def create_final_table_with_stats(final_table):
+def create_final_table_and_summary_stats(final_table):
     """
-    Create the final table with EM1-EM4 and aggregate scores, and add summary statistics at the bottom.
+    Create the final table with EM1-EM4 and aggregate scores, and display the summary statistics separately.
     """
     # Calculate summary statistics for EM1-EM4 and Aggregate
-    summary_stats = final_table[['EM1', 'EM2', 'EM3', 'EM4', 'Aggregate_EM_Score']].agg(['mean', 'median', 'std', 'min', 'max']).round(3)
+    summary_stats = final_table[['EM1', 'EM2', 'EM3', 'EM4', 'Aggregate_EM_Score']].agg(['mean', 'median', 'std', 'min', 'max']).round(1)
     
-    # Append summary statistics to the table
-    final_table_with_stats = pd.concat([final_table, summary_stats])
-
-    # Rename the index for clarity in summary statistics
-    final_table_with_stats = final_table_with_stats.rename(index={0: 'Mean', 1: 'Median', 2: 'Standard Deviation', 3: 'Min', 4: 'Max'})
-
-    print("\nFinal Combined Table with Summary Statistics:")
-    print(final_table_with_stats)
-    return final_table_with_stats
+    # Display the final table without summary statistics
+    print("\nFinal Combined Table (Main Table):")
+    print(final_table)
+    
+    # Display the summary statistics separately
+    print("\nSummary Statistics:")
+    print(summary_stats)
+    
+    return final_table, summary_stats
 
 def save_results(country_em1, summary_stats_em1, country_em2, summary_stats_em2, country_em3, summary_stats_em3, country_em4, summary_stats_em4, aggregate_em):
     """
     Save the EM1, EM2, EM3, EM4, and aggregate earnings management results to a pickle file and print them.
     """
-    final_table_with_stats = create_final_table_with_stats(aggregate_em)
+    # Correctly calculate the summary statistics for each measure
+    summary_stats = {
+        'EM1': summary_stats_em1,
+        'EM2': summary_stats_em2,
+        'EM3': summary_stats_em3,
+        'EM4': summary_stats_em4,
+    }
+    
+    summary_stats_df = pd.DataFrame(summary_stats)
 
     with open('output/em_results.pickle', 'wb') as f:
[…]

### 2024-08-11 doc/paper.qmd
deleted test EM1,2

@@
 
+# Load the results from the existing pickle file for table_1
 with open('../data/generated/table_1.pickle', 'rb') as f:
     results = pickle.load(f)
+
+# Load the results from the new pickle file for final_table and summary_stats
+with open('../output/em_results.pickle', 'rb') as f:
+    em_results = pickle.load(f)
 
@@
 The discussion on replication findings is presented in @sec-results.
-
-## delete from fikir.
 ```{python}
+#| label: final_table 
 #| echo: false
@@
 print(prep_latex_table(
-    results['replication_results'],
-    caption='Replicated table - P/B Ratios And Future Residual Income',
-    label='tab:replicated_penman'
+    em_results['final_table'],
+    caption='Replicated table - EM1, EM2, EM3, EM4, and Aggregate EM Score',
+    label='tab:final_table'
+))
+```
+
+```{python}
+#| label: summary_stats 
+#| echo: false
+#| output: asis
+
+print(prep_latex_table(
+    em_results['summary_stats'],
+    caption='Replicated table - Summary Statistics for EM1, EM2, EM3, and EM4',
+    label='tab:summary_stats'
 ))
@@
 This excludes data reported on a current, daily, or quarterly basis.
-After retaining only relevant company data, filtering out financial institutions, and focusing on the specified countries, the processed data was saved to a CSV file in the `data/pulled` directory, as specified in the configuration file `config/pull_data_cfg`​.
+After retaining only relevant company data, filtering out financial institutions, and focusing on the specified countries, the processed data was saved to a CSV file in the `data/pulled` directory, as specified in configuration file `config/pull_data_cfg`​.
 
@@
 1.	Compute Accruals by using changes in various components of a firm’s balance sheet and income statement, such as total current assets, cash, current liabilities, short-term debt, taxes payable, and depreciation:
+
 $$
@@
 $$
-\begin{aligned}
-\text{EM4}&= \frac{\text{Small Profits}}{\text{Small Losses}} \\
-& \\
-\text{where:} \\
-\text{Small Profits} & : \frac{\text{Net Earnings}_{it}}{\text{Total Assets}_{it-1}} \in [0, 0.01] \\
-\text{Small Losses}  & : \frac{\text{Net Earnings}_{it}}{\text{Total Assets}_{it-1}} \in [-0.01, 0) \\
-\end{aligned}
+\text{EM4} = \frac{\text{Small Profits}}{\text{Small Losses}}
+$$
+$$
+\begin{alignedat}{2}
+\text{where:} \\ \quad & \text{Small Profits} && = \frac{\text{Net Earnings}_{it}}{\text{Total Assets}_{it-1}} \in [0, 0.01] \\
+& \text{Small Losses} && = \frac{\text{Net Earnings}_{it}}{\text{Total Assets}_{it-1}} \in [-0.01, 0)
+\end{alignedat}
 $$

### 2024-08-11 code/python/do_analysis.py

@@ -18,11 +18,14 @@ def main():
     # Calculate EM4
     country_em4, summary_stats_em4 = calculate_em4(financial_data)
 
-    # Calculate Aggregate Earnings Management Score
-    aggregate_em = calculate_aggregate(country_em1, country_em2, country_em3, country_em4)
+    # Calculate Aggregate Earnings Management Score and create final table
+    final_table = calculate_aggregate(country_em1, country_em2, country_em3, country_em4)
 
-    # Save and print results
-    save_results(country_em1, summary_stats_em1, country_em2, summary_stats_em2, country_em3, summary_stats_em3, country_em4, summary_stats_em4, aggregate_em)
+    # Create and display the final table and summary statistics
+    final_table, summary_stats = create_final_table_and_summary_stats(final_table)
+
+    # Save only the final table and summary statistics
+    save_results(final_table, summary_stats)
 
 def load_data():
     """
@@ -179,8 +182,8 @@ def create_final_table_and_summary_stats(final_table):
     """
     Create the final table with EM1-EM4 and aggregate scores, and display the summary statistics separately.
     """
-    # Calculate summary statistics for EM1-EM4 and Aggregate
-    summary_stats = final_table[['EM1', 'EM2', 'EM3', 'EM4', 'Aggregate_EM_Score']].agg(['mean', 'median', 'std', 'min', 'max']).round(1)
+    # Calculate summary statistics for EM1-EM4 
+    summary_stats = final_table[['EM1', 'EM2', 'EM3', 'EM4']].agg(['mean', 'median', 'std', 'min', 'max']).round(3)
     
     # Display the final table without summary statistics
     print("\nFinal Combined Table (Main Table):")
@@ -192,62 +195,15 @@ def create_final_table_and_summary_stats(final_table):
     
     return final_table, summary_stats
 
-def save_results(country_em1, summary_stats_em1, country_em2, summary_stats_em2, country_em3, summary_stats_em3, country_em4, summary_stats_em4, aggregate_em):
+def save_results(final_table, summary_stats):
     """
-    Save the EM1, EM2, EM3, EM4, and aggregate earnings management results to a pickle file and print them.
+    Save the final table and summary statistics to a pickle file and print them.
     """
-    # Correctly calculate the summary statistics for each measure
-    summary_stats = {
-        'EM1': summary_stats_em1,
-        'EM2': summary_stats_em2,
[…]

### 2024-08-11 code/python/testEM1.py

@@ -1,132 +1,0 @@
-# We start by loading the libraries that we will use in this analysis.
-import pickle
-import pandas as pd
-
-def main():
-    # Load prepared data
-    financial_data = load_data()
-
-    # Calculate accruals
-    financial_data = calculate_accruals(financial_data)
-
-    # Calculate Operating Cash Flow (CFO)
-    financial_data = calculate_cfo(financial_data)
-
-    # Calculate EM1
-    em1_results = calculate_em1(financial_data)
-
-    # Save and print results
-    save_results(em1_results)
-
-def load_data():
-    """
-    Load the prepared financial data from the previous step.
-    """
-    df = pd.read_csv('data/generated/financial_data_prepared.csv')
-    return df
-
-def calculate_accruals(df):
-    """
-    Calculate accruals using the provided formula.
-    If a firm does not report information on taxes payable or short-term debt,
-    then the change in both variables is assumed to be zero.
-    """
-
-    # Group by unique firm (item6105) and calculate diffs within each firm
-    df['delta_CA'] = df.groupby('item6105')['item2201'].diff()  # Change in total current assets
-    df['delta_Cash'] = df.groupby('item6105')['item2003'].diff()  # Change in cash and cash equivalents
-    df['delta_CL'] = df.groupby('item6105')['item3101'].diff()  # Change in total current liabilities
-    df['delta_STD'] = df.groupby('item6105')['item3051'].diff().fillna(0)  # Apply fillna only for STD
-    df['delta_TP'] = df.groupby('item6105')['item3063'].diff().fillna(0)  # Apply fillna only for TP
-    df['Dep'] = df['item1151']  # No fillna applied to Depreciation and amortization expense
-
-    df['Accruals'] = (df['delta_CA'] - df['delta_Cash']) - (df['delta_CL'] - df['delta_STD'] - df['delta_TP']) - df['Dep']
-
-    # Identify rows with NaN in Accruals
-    nan_accruals = df[df['Accruals'].isna()]
-
-    # Print the first 10 rows with NaN in Accruals
-    print("First 10 rows with NaN in Accruals:")
-    print(nan_accruals.head(10))
-
-    # Count and print the number of NaN rows in Accruals
-    nan_count = df['Accruals'].isna().sum()
-    print(f"Number of rows with NaN in Accruals: {nan_count}")
-
-    # Print the first few rows to check the calculated accruals
-    print(df[['item6105', 'year_', 'Accruals']].head(10))
[…]

### 2024-08-11 code/python/testEM2.py

@@ -1,134 +1,0 @@
-# We start by loading the libraries that we will use in this analysis.
-import pickle
-import pandas as pd
-
-def main():
-    # Load the prepared data
-    financial_data = load_data()
-
-    # Calculate EM1
-    country_em1, summary_stats_em1 = calculate_em1(financial_data)
-
-    # Calculate EM2
-    country_em2, summary_stats_em2 = calculate_em2(financial_data)
-
-    # Save and print results
-    save_results(country_em1, summary_stats_em1, country_em2, summary_stats_em2)
-
-def load_data():
-    """
-    Load the prepared financial data from the previous step.
-    """
-    df = pd.read_csv('data/generated/financial_data_prepared.csv')
-    return df
-
-def calculate_em1(df):
-    """
-    Calculate EM1 for each firm and then take the country-level median.
-    This function combines all necessary steps into one.
-    """
-    # Step 1: Calculate Accruals
-    # Group by unique firm (item6105) and calculate diffs within each firm
-    # If a firm does not report information on taxes payable or short-term debt,
-    # then the change in both variables is assumed to be zero (per paper).
-    df['delta_CA'] = df.groupby('item6105')['item2201'].diff()  # Change in total current assets
-    df['delta_Cash'] = df.groupby('item6105')['item2003'].diff()  # Change in cash and cash equivalents
-    df['delta_CL'] = df.groupby('item6105')['item3101'].diff()  # Change in total current liabilities
-    df['delta_STD'] = df.groupby('item6105')['item3051'].diff().fillna(0)  # Apply fillna only for STD
-    df['delta_TP'] = df.groupby('item6105')['item3063'].diff().fillna(0)  # Apply fillna only for TP
-    df['Dep'] = df['item1151']  # Depreciation and amortization expense
-
-    df['Accruals'] = (df['delta_CA'] - df['delta_Cash']) - (df['delta_CL'] - df['delta_STD'] - df['delta_TP']) - df['Dep']
-
-    # Step 2: Calculate Operating Cash Flow (CFO) by subtracting accruals from operating income.
-    # Handle NaN values where necessary.
-    df['CFO'] = df['item1250'] - df['Accruals']
-
-    # Step 3: Calculate the standard deviations for Operating Income and CFO for each firm (grouped by item1605)
[…]

### 2024-08-11 code/python/do_analysis.py
combine summary and final table

@@ -21,11 +21,11 @@ def main():
     # Calculate Aggregate Earnings Management Score and create final table
     final_table = calculate_aggregate(country_em1, country_em2, country_em3, country_em4)
 
-    # Create and display the final table and summary statistics
-    final_table, summary_stats = create_final_table_and_summary_stats(final_table)
+    # Create the final combined table with both metrics and summary statistics
+    final_combined_table = create_final_combined_table(final_table)
 
-    # Save only the final table and summary statistics
-    save_results(final_table, summary_stats)
+    # Save the final combined table
+    save_results(final_combined_table)
 
 def load_data():
     """
@@ -195,14 +195,45 @@ def create_final_table_and_summary_stats(final_table):
     
     return final_table, summary_stats
 
-def save_results(final_table, summary_stats):
+def create_final_combined_table(final_table):
     """
-    Save the final table and summary statistics to a pickle file and print them.
+    Create a final combined table with EM1-EM4, aggregate scores, and summary statistics.
+    """
+    # Calculate summary statistics for EM1-EM4
+    summary_stats = final_table[['EM1', 'EM2', 'EM3', 'EM4']].agg(['mean', 'median', 'std', 'min', 'max']).round(3)
+    
+    # Rename the index of summary_stats to be more descriptive
+    summary_stats.index = ['Mean', 'Median', 'Std', 'Min', 'Max']
+    
+    # Create a DataFrame for the summary stats with the same columns as final_table
+    summary_stats_df = pd.DataFrame(
+        summary_stats.values,
+        columns=['EM1', 'EM2', 'EM3', 'EM4'],
+        index=['Mean', 'Median', 'Std', 'Min', 'Max']
+    ).reset_index()
+    
+    # Add the names in the 'item6026' column (which originally holds the country names)
+    summary_stats_df['item6026'] = summary_stats_df['index']
+    summary_stats_df['Aggregate_EM_Score'] = ''
+    
+    # Reorder columns to match the final_table structure and drop the 'index' column
+    summary_stats_df = summary_stats_df[['item6026', 'EM1', 'EM2', 'EM3', 'EM4', 'Aggregate_EM_Score']]
+    
+    # Append the summary stats to the final table
+    final_combined_table = pd.concat([final_table, summary_stats_df], ignore_index=True)
+    
+    print("\nFinal Combined Table (with Summary Statistics):")
+    print(final_combined_table)
+    
+    return final_combined_table
+
+def save_results(final_combined_table):
+    """
[…]

### 2024-08-11 doc/paper.qmd
finalise replicated table

@@
 ```{python}
-#| label: final_table 
+#| label: final_combined_table
 #| echo: false
@@
 print(prep_latex_table(
-    em_results['final_table'],
-    caption='Replicated table - EM1, EM2, EM3, EM4, and Aggregate EM Score',
-    label='tab:final_table'
-))
-```
-
-```{python}
-#| label: summary_stats 
-#| echo: false
-#| output: asis
-
-print(prep_latex_table(
-    em_results['summary_stats'],
-    caption='Replicated table - Summary Statistics for EM1, EM2, EM3, and EM4',
-    label='tab:summary_stats'
+    em_results['final_combined_table'], 
+    caption='Replicated table - EM1, EM2, EM3, EM4, Aggregate EM Score, and Summary Statistics',
+    label='tab:final_combined_table'
 ))

### 2024-08-11 code/python/do_analysis.py

@@ -185,14 +185,6 @@ def create_final_table_and_summary_stats(final_table):
     # Calculate summary statistics for EM1-EM4 
     summary_stats = final_table[['EM1', 'EM2', 'EM3', 'EM4']].agg(['mean', 'median', 'std', 'min', 'max']).round(3)
     
-    # Display the final table without summary statistics
-    print("\nFinal Combined Table (Main Table):")
-    print(final_table)
-    
-    # Display the summary statistics separately
-    print("\nSummary Statistics:")
-    print(summary_stats)
-    
     return final_table, summary_stats
 
 def create_final_combined_table(final_table):
@@ -219,8 +211,18 @@ def create_final_combined_table(final_table):
     # Reorder columns to match the final_table structure and drop the 'index' column
     summary_stats_df = summary_stats_df[['item6026', 'EM1', 'EM2', 'EM3', 'EM4', 'Aggregate_EM_Score']]
     
+    # Format the summary_stats_df to include trailing zeros for all numbers
+    summary_stats_df[['EM1', 'EM2', 'EM3', 'EM4']] = summary_stats_df[['EM1', 'EM2', 'EM3', 'EM4']].applymap(lambda x: f"{x:.3f}")
+    
+    # Insert an empty row after index 30
+    empty_row = pd.DataFrame([['', '', '', '', '', '']], columns=final_table.columns)
+    final_table_with_blank = pd.concat([final_table.iloc[:31], empty_row, final_table.iloc[31:]], ignore_index=True)
+    
+    # Format the final_table_with_blank to include trailing zeros for all numbers
+    final_table_with_blank[['EM1', 'EM2', 'EM3', 'EM4']] = final_table_with_blank[['EM1', 'EM2', 'EM3', 'EM4']].applymap(lambda x: f"{float(x):.3f}" if x != '' else x)
+    
     # Append the summary stats to the final table
-    final_combined_table = pd.concat([final_table, summary_stats_df], ignore_index=True)
+    final_combined_table = pd.concat([final_table_with_blank, summary_stats_df], ignore_index=True)
     
     print("\nFinal Combined Table (with Summary Statistics):")
     print(final_combined_table)

### 2024-08-11 code/python/do_analysis.py
update table 1 replication (blank row)

@@ -221,6 +221,9 @@ def create_final_combined_table(final_table):
     # Format the final_table_with_blank to include trailing zeros for all numbers
     final_table_with_blank[['EM1', 'EM2', 'EM3', 'EM4']] = final_table_with_blank[['EM1', 'EM2', 'EM3', 'EM4']].applymap(lambda x: f"{float(x):.3f}" if x != '' else x)
     
+    # Ensure Aggregate_EM_Score is rounded to one decimal place and formatted with one decimal
+    final_table_with_blank['Aggregate_EM_Score'] = final_table_with_blank['Aggregate_EM_Score'].apply(lambda x: f"{float(x):.1f}" if x != '' else x)
+    
     # Append the summary stats to the final table
     final_combined_table = pd.concat([final_table_with_blank, summary_stats_df], ignore_index=True)
     

### 2024-08-11 code/python/prepare_data.py

@@ -9,7 +9,6 @@ import pickle
 from utils import read_config, setup_logging
 
 log = setup_logging()
-
 def main():
     log.info("Preparing data for analysis ...")
     cfg = read_config('config/prepare_data_cfg.yaml')
@@ -75,9 +74,13 @@ def main():
     # Remove decimal places
     table_1['# Firm-years'] = table_1['# Firm-years'].apply(lambda x: int(round(x)))
     
+    # Insert an empty row after index 30
+    empty_row = pd.DataFrame([['', '']], columns=table_1.columns)
+    table_1_with_blank = pd.concat([table_1.iloc[:31], empty_row, table_1.iloc[31:]], ignore_index=True)
+    
     # Save Table 1 to a pickle file
     results = {
-        "table_1": table_1
+        "table_1": table_1_with_blank
     }
     with open(cfg['table_1_save_path'], 'wb') as f:
         pickle.dump(results, f)

### 2024-08-11 doc/paper.qmd
update latex tables

@@
 def prep_latex_table(df, caption=None, label=None, fontsize='\\fontsize{11}{13}\\selectfont'):
-    df = escape_special_characters(df.reset_index())
+    df = escape_special_characters(df.reset_index(drop=True))  # Drop the index column from output tables for better readability
     num_columns = len(df.columns)
@@
     em_results['final_combined_table'], 
-    caption='Replicated table - EM1, EM2, EM3, EM4, Aggregate EM Score, and Summary Statistics',
+    caption='Replicated table - Country scores for earnings management measures (Sorted by aggregate earnings management)',
     label='tab:final_combined_table'
@@
     results['table_1'],
-    caption='Replicated table - Number of Firm-Year Observations per Country',
+    caption='Replicated table - Descriptive statistics: Number of firm-year observations per country',
     label='tab:table_1'

### 2024-08-11 doc/paper.qmd
table formatting

@@
 
+# Rename the 'item6026' and 'aggregate_EM_Score' columns
+em_results['final_combined_table'].rename(columns={
+    'item6026': 'Country',
+    'Aggregate_EM_Score': 'Aggregate EM score'
+}, inplace=True)
+
+# Render the final table with the updated column names
 print(prep_latex_table(

### 2024-08-11 doc/paper.qmd
latex formatting

--- 2024-08-11 doc/paper.qmd
@@
 
-def prep_latex_table(df, caption=None, label=None, fontsize='\\fontsize{11}{13}\\selectfont'):
+def prep_latex_table(df, caption=None, label=None, fontsize='\\fontsize{11}{13}\\selectfont', description=None):
     df = escape_special_characters(df.reset_index(drop=True))  # Drop the index column from output tables for better readability
@@
     column_format = 'l' + 'r' * (num_columns - 1)
+
     latex_table = df.to_latex(
@@
         float_format=custom_float_format,
+        escape=False  # Important to disable escaping so LaTeX commands in column names are respected
     )
+
+    # Construct description to the table
+    full_caption = f"{caption}\\\\\n\\textnormal{{{description}}}" if description else caption
+
     latex_table_lines = [
@@
         latex_table,
-        f"\\caption{{{caption}}}" if caption else "",
+        f"\\caption{{{full_caption}}}" if full_caption else "",
         f"\\label{{{label}}}" if label else "",
@@
 
+# Rename the columns to ensure consistency with paper
+em_results['final_combined_table'].rename(columns={
+    'item6026': 'Country',
+    'EM1': 'EM1 (−)',
+    'EM2': 'EM2 (−)',
+    'EM3': 'EM3 (+)',
+    'EM4': 'EM4 (+)',
+    'Aggregate_EM_Score': 'Aggregate EM score'
+}, inplace=True)
+
 print(prep_latex_table(
@@
     caption='Replicated table - Country scores for earnings management measures (Sorted by aggregate earnings management)',
-    label='tab:final_combined_table'
+    label='tab:final_combined_table',
+    description = (
+    "\\fontsize{10pt}{12pt}\\selectfont " 
+    "The variables are computed from 123,469 firm-year observations for fiscal years 1990 to 1999 across 31 countries and 18,040 non-financial firms.
+Data are obtained from the Worldscope Database (July 2024 version).
+The definitions and calculations of EM1-4, as well as the aggregate score, are thoroughly presented in Section 4.
+The sign in the column heading indicates whether higher scores for the respective EM measure imply more earnings management (+) or less earnings management (−)."
+    )
+
 ))

### 2024-08-11 code/python/do_analysis.py
update config/do_analysis_cfg.yaml

@@ -1,10 +1,19 @@
 # We start by loading the libraries that we will use in this analysis.
 import pickle
 import pandas as pd
+from utils import read_config, setup_logging
+
+# Set up logging
+log = setup_logging()
 
 def main():
-    # Load the prepared data
-    financial_data = load_data()
+    log.info("Performing main analysis...")
+    
+    # Load the configuration file
+    cfg = read_config('config/do_analysis_cfg.yaml')
+    
+    # Load the prepared financial data using the path from the config file
+    financial_data = load_data(cfg['prepared_data_save_path'])
 
     # Calculate EM1
     country_em1, summary_stats_em1 = calculate_em1(financial_data)
@@ -24,14 +33,16 @@ def main():
     # Create the final combined table with both metrics and summary statistics
     final_combined_table = create_final_combined_table(final_table)
 
-    # Save the final combined table
-    save_results(final_combined_table)
+    # Save the final combined table using the path from the config file
+    save_results(final_combined_table, cfg['results'])
 
-def load_data():
+    log.info("Performing main analysis...Done!")
+
+def load_data(data_path):
     """
-    Load the prepared financial data from the previous step.
+    Load the prepared financial data from the specified path.
     """
-    df = pd.read_csv('data/generated/financial_data_prepared.csv')
+    df = pd.read_csv(data_path)
     return df
 
 def calculate_em1(df):
@@ -232,11 +243,11 @@ def create_final_combined_table(final_table):
     
     return final_combined_table
 
-def save_results(final_combined_table):
+def save_results(final_combined_table, save_path):
     """
-    Save the final combined table to a pickle file.
+    Save the final combined table to a pickle file at the specified path.
     """
-    with open('output/em_results.pickle', 'wb') as f:
+    with open(save_path, 'wb') as f:
         pickle.dump({
             'final_combined_table': final_combined_table
         }, f)

### 2024-08-11 doc/paper.qmd
EM4 calc. update: at lest 5 obs. small profits

@@
 However, the original paper might have included additional data cleaning steps not explicitly mentioned, such as handling outliers, specific industry exclusions, or other criteria, which could affect the final counts.
-## table 2 correctly referenced?
+
 Given that the number of observations in this project is significantly higher than in the original study, Table 1 from @Leuz_2003 was partially replicated (focusing only on columns related to countries and firm-year observations) to highlight specific discrepancies that could arise for certain countries.
-The results in the Table 2 below show a higher number of firm-year observations for almost all countries compared to the paper by @Leuz_2003, likely due to database updates leading to a more comprehensive dataset.
+So, the results in the Table 2 below show a higher number of firm-year observations for almost all countries compared to the paper by @Leuz_2003, likely due to database updates leading to a more comprehensive dataset.
 The trend in the number of firm-year observations across countries is consistent with the original study, with larger capital markets like Japan, the United Kingdom, and the United States having the highest numbers.

### 2024-08-11 code/python/do_analysis.py

@@ -143,6 +143,7 @@ def calculate_em4(df):
     """
     Calculate EM4, which is the ratio of the number of small profits to the number of small losses for each country.
     Small profits and losses are defined based on net earnings (item1651) scaled by lagged total assets.
+    Only include countries with at least 5 small losses.
     """
     # Step 1: Calculate Net Earnings scaled by Lagged Total Assets
     df['scaled_net_earnings'] = df['item1651'] / df['lagged_total_assets']
@@ -151,16 +152,26 @@ def calculate_em4(df):
     df['small_profits'] = ((df['scaled_net_earnings'] >= 0) & (df['scaled_net_earnings'] <= 0.01)).astype(int)
     df['small_losses'] = ((df['scaled_net_earnings'] >= -0.01) & (df['scaled_net_earnings'] < 0)).astype(int)
 
-    # Step 3: Calculate EM4 as the ratio of Small Profits to Small Losses for each country
-    country_em4 = df.groupby('item6026').apply(
+    # Step 3: Filter countries with at least 5 small losses
+    country_counts = df.groupby('item6026')['small_losses'].sum()
+    eligible_countries = country_counts[country_counts >= 5].index
+
+    filtered_df = df[df['item6026'].isin(eligible_countries)]
+
+    # Step 4: Calculate EM4 as the ratio of Small Profits to Small Losses for each eligible country
+    country_em4 = filtered_df.groupby('item6026').apply(
         lambda x: x['small_profits'].sum() / max(1, x['small_losses'].sum())  # Avoid division by zero
     ).reset_index()
     country_em4.columns = ['item6026', 'EM4']
     # Round the EM4 results to three decimal places
     country_em4['EM4'] = country_em4['EM4'].round(3)
 
-    # Step 4: Calculate summary statistics (mean, median, std, min, max) for EM4 across countries
+    # Step 5: Calculate summary statistics (mean, median, std, min, max) for EM4 across countries
     summary_stats_em4 = country_em4['EM4'].agg(['mean', 'median', 'std', 'min', 'max']).round(3)
+
+    # Print how many countries were excluded
+    excluded_countries = len(df['item6026'].unique()) - len(eligible_countries)
+    print(f"{excluded_countries} countries were excluded due to having fewer than 5 small losses.")
 
     return country_em4, summary_stats_em4
 

### 2024-08-11 doc/paper.qmd
again update paper.qmd

@@
 Note that I used `item1651` (Net Income before Preferred Dividends) as the variable for net earnings, as mentioned in @sec-step2_preparation.
+I also ensured that each country included in the EM4 calculation had at least five small loss observations, as per the methodology outlined by @Leuz_2003.
+The review found that no countries were excluded from the analysis due to this criterion.
 
@@
 In the future, this repository can be cloned or forked (if made public) to kickstart further projects on earnings management measures analysis.
+Additionally, the included presentation template can be rendered alongside the paper for further academic use, providing a visually engaging summary of the research findings.
 Thanks for reading!

### 2024-08-11 doc/presentation.qmd
rendering presentation

---
title: |
  | Earnings Management and Investor Protection:
  | Accounting Reading Group - Assignment III
author: "Melisa Mazaeva"
institute: "Humboldt-Universität zu Berlin – Summer 2024"
date: today
fontsize: "9pt"
pdf-engine: xelatex # pdflatex creates rastered fonts
format: 
  beamer:
    slide-level: 3
    number-sections: true
    toc: false
header-includes:
- \usepackage{booktabs} 
- \usepackage{threeparttable}
- \usepackage{graphicx}
- \usepackage{etoolbox}
- \AtBeginEnvironment{verbatim}{\fontsize{7}{7}\selectfont}  # This sets the font size for verbatim environments
- \input{beamer_theme_trr266.sty}
bibliography: references.bib
biblio-style: apsr
---

```{python}
#| echo: false
#| output: false

import pickle

with open('../output/results.pickle', 'rb') as f:
        results = pickle.load(f)

min_fyear, max_fyear, no_unique_firms = results['Desc information'].values()

def rename_var_to_label(table, italic=True):
    for var_name, var_label in results['Variable Names'].items():
        var_label = f'\\textit{{{var_label}}}' if italic else var_label
        table = table.replace(var_name, var_label)
    return table
```

\pagebreak

# Motivation

### Discretionary Accruals

- Are used in tons of papers
- As measurement constructs for almost everything
  - Earnings Management
  - Earnings Quality
  - Audit Quality
  - Management Quality
- But how do they look like?

### Data

```{python}
#| echo: false
#| output: asis

lat_tab = rename_var_to_label(results['Descriptive Statistics']).split('\n')
# lat_tab[4] = '\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]'
# lat_tab[6] = '\\hline\\\\[-1.8ex]'
# lat_tab[-4] = '\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]'

print('\n'.join([
    '\\begin{table}[!htbp]',  
    '\\resizebox{\\textheight}{!}{',
    '\\begin{threeparttable}',
    '\n'.join(lat_tab[3:-2]),
    '\\begin{tablenotes}[flushleft]',
    '\setlength{\labelsep}{0pt}',
    f'\\item Note: The data is obtained from the Compustat U.S. as provided by WRDS. The sample covers the period {int(min_fyear)} to {int(max_fyear)} and {no_unique_firms:,} unique firms.',
    '\end{tablenotes}',
    '\end{threeparttable}',
    '}',
    '\\end{table}',
]))
```

### Distribution of Discretionary Accruals over time - unrestricted sample

```{python}
#| label: fig-boxplot_full
#| fig-cap: "Distribution of Discretionary Accruals over time - unrestricted sample"
#| echo: false
results['fig_boxplot_full'].set_figheight(2.6)
results['fig_boxplot_full']
```

### Distribution of Discretionary Accruals over time - analysis sample

```{python}
#| label: fig-boxplot_smp
#| fig-cap: "Distribution of Discretionary Accruals over Time - analysis sample"
#| echo: false
results['fig_boxplot_smp'].set_figheight(2.6)
results['fig_boxplot_smp']
```

### Correlation of Modified Jones with Dechow and Dichev DA

```{python}
#| label: fig-scatter_md_dd
#| fig-cap: "Correlation of Modified Jones with Dechow and Dichev DA"
#| echo: false
results['fig_scatter_md_dd'].set_figheight(3)
results['fig_scatter_md_dd']
```

### Association of Dechow and Dichev DA with firm size

```{python}
#| label: fig-scatter_dd_lnta
#| fig-cap: "Dechow and Dichev DA and Firm Size"
#| echo: false

results['fig_scatter_dd_lnta'].set_figheight(3)
results['fig_scatter_dd_lnta']
```

### No presentation without correlation table...

```{python}
#| echo: false
#| output: asis

lat_tab = rename_var_to_label(results['Correlations'], False).split('\n')
lat_tab[2] = '\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]'
lat_tab[4] = '\\hline\\\\[-1.8ex]'
lat_tab[-7] = '\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]'

print('\n'.join([
    '\\begin{center}',
    '\\resizebox*{\\textwidth}{!}{',
    '\\begin{tabular}{lccccccc}',
    '\n'.join(lat_tab[2:-6]),
    '\\end{tabular}}',
    '\\end{center}'
]))
```

### ... and regression table

```{python}
#| echo: false
#| output: asis

print('\n'.join([
    '\\begin{center}',
    '\\resizebox*{!}{0.8\\textheight}{', 
    '\\begin{tabular}{lcc}',
    rename_var_to_label('\n'.join(results['Regression'].split('\n')[4:-2])),
    '\\end{tabular}}',
    '\\end{center}'
]))
```

### Conclusion

- Discretionary accruals are simply wonderful! 
- With these distributional properties and the power of large samples they will almost always be significant in one direction or the other regardless the setting or controls
- Use them for every economic construct under the sun!

### 2024-08-11 doc/presentation.qmd
presentation workflow

@@
 import pickle
+import pandas as pd
 
-with open('../output/results.pickle', 'rb') as f:
+# Load the results from the existing pickle file for table_1
+with open('../data/generated/table_1.pickle', 'rb') as f:
     results = pickle.load(f)
 
-min_fyear, max_fyear, no_unique_firms = results['Desc information'].values()
+# Load the results from the new pickle file for final_table and summary_stats
+with open('../output/em_results.pickle', 'rb') as f:
+    em_results = pickle.load(f)
 
-def rename_var_to_label(table, italic=True):
-    for var_name, var_label in results['Variable Names'].items():
-        var_label = f'\\textit{{{var_label}}}' if italic else var_label
-        table = table.replace(var_name, var_label)
-    return table
+def escape_special_characters(df):
+    df.columns = [col.replace('_', '\\_').replace('#', '\\#') for col in df.columns]
+    return df
+
+# Custom formatter function
+def custom_float_format(x):
+    return f"{x:.3f}".lstrip('0') if pd.notnull(x) else ""
+
+def prep_latex_table(df, caption=None, label=None, fontsize='\\fontsize{11}{13}\\selectfont', description=None):
+    df = escape_special_characters(df.reset_index(drop=True))  # Drop the index column from output tables for better readability
+    num_columns = len(df.columns)
+    column_format = 'l' + 'r' * (num_columns - 1)
+
+    latex_table = df.to_latex(
+        column_format=column_format,
+        index=False, 
+        float_format=custom_float_format,
+        escape=False  # Important to disable escaping so LaTeX commands in column names are respected
+    )
+
+    # Construct description to the table
+    full_caption = f"{caption}\\\\\n\\textnormal{{{description}}}" if description else caption
+
+    latex_table_lines = [
+        "\\begin{table}[htbp]",
+        "\\centering",
+        fontsize,  # Set font size for the table
+        latex_table,
+        f"\\caption{{{full_caption}}}" if full_caption else "",
+        f"\\label{{{label}}}" if label else "",
+        "\\end{table}"
+    ]
+    return "\n".join(line for line in latex_table_lines if line)
 ```
-
-\pagebreak
 
@@
 
-### Discretionary Accruals
+## Understanding Systematic Differences in Earnings Management across Countries
+
+- The paper of @Leuz_2003 focuses on the relationship between investor protection and earnings management.
+- Investor protection plays a crucial role in limiting the extent of earnings management by reducing the opportunities for managers to manipulate earnings.
+- Analyzing the relationship between EM and investor protection across different countries provides insights into how legal and institutional frameworks impact financial reporting practices.
+- Earnings management (EM) measures reflect the extent to which managers manipulate financial reports to meet certain objectives, such as meeting earnings targets or influencing stock prices.
+
+## Project Objective
+
+- The primary objective is to demonstrate a reproducible and collaborative research workflow by calculating and analyzing EM measures across different countries.
+- The study aims to replicate and extend the findings of @Leuz_2003, exploring the relationship between EM and investor protection using Worldscope datasets for the time range 1990 to 1999.
+- The analysis provides insights into the levels of earnings management across countries, comparing them with the original findings of Leuz, Nanda, and Wysocki (2003).
+
+## Project Relevance
+
+- Understanding EM trends across countries aids in assessing global financial transparency.
+- Promotes transparency and reproducibility in empirical accounting research.
+
+# Research Design
+
+- Data sourced from the Worldscope Database (1990-1999) July 2024 version.
+- Replication includes calculation of **four earnings management measures**:
+  1. Earnings smoothing using accruals (EM1)
+  2. Correlation between changes in accruals and cash flows (EM2)
+  3. Magnitude of accruals (EM3)
+  4. Ratio of small profits to small losses (EM4)
+
+# Assumptions
+
+- **Database Version**: Analysis uses the July 2024 Worldscope data, which may differ from the November 2000 version due to updates and restatements.
+- **Key Terms**: Earnings management (EM) and investor protection defined as per @Leuz_2003 for consistency.
+- **Handling Negative Values**: Negative values in key metrics (e.g., operating income) were included to capture the full scope of EM activities.
+- **Variable Selection**: Chose specific Worldscope variables (e.g., item1151 over item4051) based on best practices; differences may affect results.
+- **Currency Consistency**: EM measures are based on scaled variables; currency consistency ensures comparability across countries.
+
+### Discretionary Accruals ##delete
 
@@
 
-### Data
+### Data - Replication Steps
 
+Step 1: Pulling the Data and Managing the Databases
+
+Step 2: Data Preparation
+
++  To verify the robustness of descriptive statistics before the main replication, Table 1 from Leuz et al. was replicated, yielding results that closely align with the original study:
 ```{python}
+#| label: table_1 
 #| echo: false
@@
 
-lat_tab = rename_var_to_label(results['Descriptive Statistics']).split('\n')
-# lat_tab[4] = '\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]'
-# lat_tab[6] = '\\hline\\\\[-1.8ex]'
-# lat_tab[-4] = '\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]'
-
-print('\n'.join([
-    '\\begin{table}[!htbp]',  
-    '\\resizebox{\\textheight}{!}{',
-    '\\begin{threeparttable}',
-    '\n'.join(lat_tab[3:-2]),
-    '\\begin{tablenotes}[flushleft]',
-    '\setlength{\labelsep}{0pt}',
-    f'\\item Note: The data is obtained from the Compustat U.S. as provided by WRDS. The sample covers the period {int(min_fyear)} to {int(max_fyear)} and {no_unique_firms:,} unique firms.',
-    '\end{tablenotes}',
-    '\end{threeparttable}',
-    '}',
-    '\\end{table}',
-]))
-```
-
-### Distribution of Discretionary Accruals over time - unrestricted sample
-
-```{python}
-#| label: fig-boxplot_full
-#| fig-cap: "Distribution of Discretionary Accruals over time - unrestricted sample"
-#| echo: false
-results['fig_boxplot_full'].set_figheight(2.6)
-results['fig_boxplot_full']
-```
-
-### Distribution of Discretionary Accruals over time - analysis sample
-
-```{python}
-#| label: fig-boxplot_smp
-#| fig-cap: "Distribution of Discretionary Accruals over Time - analysis sample"
-#| echo: false
-results['fig_boxplot_smp'].set_figheight(2.6)
-results['fig_boxplot_smp']
-```
-
-### Correlation of Modified Jones with Dechow and Dichev DA
-
-```{python}
-#| label: fig-scatter_md_dd
-#| fig-cap: "Correlation of Modified Jones with Dechow and Dichev DA"
-#| echo: false
-results['fig_scatter_md_dd'].set_figheight(3)
-results['fig_scatter_md_dd']
-```
-
-### Association of Dechow and Dichev DA with firm size
-
-```{python}
-#| label: fig-scatter_dd_lnta
-#| fig-cap: "Dechow and Dichev DA and Firm Size"
-#| echo: false
-
-results['fig_scatter_dd_lnta'].set_figheight(3)
-results['fig_scatter_dd_lnta']
-```
-
-### No presentation without correlation table...
-
-```{python}
-#| echo: false
-#| output: asis
-
-lat_tab = rename_var_to_label(results['Correlations'], False).split('\n')
-lat_tab[2] = '\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]'
-lat_tab[4] = '\\hline\\\\[-1.8ex]'
-lat_tab[-7] = '\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]'
-
-print('\n'.join([
-    '\\begin{center}',
-    '\\resizebox*{\\textwidth}{!}{',
-    '\\begin{tabular}{lccccccc}',
-    '\n'.join(lat_tab[2:-6]),
-    '\\end{tabular}}',
-    '\\end{center}'
-]))
-```
-
-### ... and regression table
-
-```{python}
-#| echo: false
-#| output: asis
-
-print('\n'.join([
-    '\\begin{center}',
-    '\\resizebox*{!}{0.8\\textheight}{', 
-    '\\begin{tabular}{lcc}',
-    rename_var_to_label('\n'.join(results['Regression'].split('\n')[4:-2])),
-    '\\end{tabular}}',
-    '\\end{center}'
-]))
+print(prep_latex_table(
+    results['table_1'],
+    caption='Replicated table - Descriptive statistics: Number of firm-year observations per country',
+    label='tab:table_1'
+))
 ```
@@
 
-- Discretionary accruals are simply wonderful! 
-- With these distributional properties and the power of large samples they will almost always be significant in one direction or the other regardless the setting or controls
-- Use them for every economic construct under the sun!
+- Significant differences in earnings management practices across countries.
+- Strong alignment with original findings in Leuz, Nanda, and Wysocki (2003).
+- **Future Work**: Repository can be cloned or forked for further analysis.
+
+# References {-}
+\setlength{\parindent}{-0.2in}
+\setlength{\leftskip}{0.2in}
+\setlength{\parskip}{8pt}
+\noindent

### 2024-08-11 doc/presentation.qmd
output table 1 in presentation

@@
 +  To verify the robustness of descriptive statistics before the main replication, Table 1 from Leuz et al. was replicated, yielding results that closely align with the original study:
+---
 ```{python}
@@
     caption='Replicated table - Descriptive statistics: Number of firm-year observations per country',
-    label='tab:table_1'
+    label='tab:table_1',
+    fontsize='\\fontsize{5pt}{6pt}\\selectfont'  # Set smaller font size
 ))

### 2024-08-11 doc/paper.qmd
references update

@@
 **IDE**: Integrated Development Environment  
+**SIC**: Standard Industrial Classification  
 **UK**: United Kingdom  
@@
 
-In line with the methodology used by @Leuz_2003, financial institutions were removed from the analysis based on their SIC codes.
+In line with the methodology used by @Leuz_2003, financial institutions were removed from the analysis based on their Standard Industrial Classification (SIC) codes.
 This was done by filtering out rows where the `item7021` identifier, representing the SIC code, falls within the range of 6000 to 6999.

### 2024-08-11 code/python/do_analysis.py
final commit!

@@ -236,17 +236,16 @@ def create_final_combined_table(final_table):
     # Format the summary_stats_df to include trailing zeros for all numbers
     summary_stats_df[['EM1', 'EM2', 'EM3', 'EM4']] = summary_stats_df[['EM1', 'EM2', 'EM3', 'EM4']].applymap(lambda x: f"{x:.3f}")
     
-    # Insert an empty row after index 30
+    # Insert an empty row after index 30 like in paper
     empty_row = pd.DataFrame([['', '', '', '', '', '']], columns=final_table.columns)
     final_table_with_blank = pd.concat([final_table.iloc[:31], empty_row, final_table.iloc[31:]], ignore_index=True)
     
     # Format the final_table_with_blank to include trailing zeros for all numbers
     final_table_with_blank[['EM1', 'EM2', 'EM3', 'EM4']] = final_table_with_blank[['EM1', 'EM2', 'EM3', 'EM4']].applymap(lambda x: f"{float(x):.3f}" if x != '' else x)
     
-    # Ensure Aggregate_EM_Score is rounded to one decimal place and formatted with one decimal
+    # Round to one decimal place like in paper
     final_table_with_blank['Aggregate_EM_Score'] = final_table_with_blank['Aggregate_EM_Score'].apply(lambda x: f"{float(x):.1f}" if x != '' else x)
     
-    # Append the summary stats to the final table
     final_combined_table = pd.concat([final_table_with_blank, summary_stats_df], ignore_index=True)
     
     print("\nFinal Combined Table (with Summary Statistics):")
